\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{float}
\usepackage[expansion=false]{microtype}
\usepackage{hyperref}
\hypersetup{hidelinks}
\usepackage{geometry}
\usepackage{orcidlink}
\geometry{margin=1in}
\setlength{\emergencystretch}{3em}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\newtheorem{remark}{Remark}
\newtheorem{definition}{Definition}

\title{A Quantitative Convergence Law for the\\
Connes--Consani--Moscovici Zeta Spectral Triple}

\author{Ronnie Andrews, Jr.\,\orcidlink{0009-0003-9724-3104}%
\thanks{Correspondence: \texttt{randrewsmath@gmail.com}}}

\date{June 2026\\[2pt]{\small Revised July 2026}}

\begin{document}
\maketitle

\begin{abstract}
We give a quantitative convergence law
for the first Riemann zero produced by the Connes--Consani--Moscovici (CCM)
zeta spectral triple: a three-budget minimum in digit space, equivalently a
single two-channel spectral edge, with every component carried at an explicit
rigor level. In detail, the CCM construction~\cite{CCM2025} and
its Connes--van Suijlekom refinement~\cite{CvS2025} produce finite real
symmetric matrices whose ground state yields an approximation $\nu_1$ to
the first
nontrivial Riemann zero $\gamma_1$ that sharpens as three parameters grow: a
prime cutoff $\lambda^2$, a basis size $N$, and the working precision $P$. The
rate of this convergence in the basis size is named the central open problem
in~\cite{CCM2025}. We characterize the convergence quantitatively, in two equivalent forms.
Intrinsically, the construction's relative error $|\nu_1-\gamma_1|/|\gamma_1|$
is controlled by a single two-channel decomposition,
\[
  E(N,\lambda^2)\le\varepsilon_{\mathrm{modes}}(N,\lambda^2)
                 +\varepsilon_{\mathrm{primes}}(\lambda^2),
\]
with both channels \emph{defined} from the construction (the finite-$N$
displacement at fixed cutoff and the cutoff displacement at infinite basis
size), a mode channel vanishing as $N\to\infty$ and a prime channel vanishing as
$\lambda^2\to\infty$, so that $\nu_1\to\gamma_1$ is exactly $E\to0$ and forces
both limits. The two channels are the finite-$N$ transition and the extreme tail
of a single spectral edge (the lower edge of the equilibrium measure of the
archimedean field, whose bulk density is the Riemann--von Mangoldt law), so
$D_{\mathrm{nModes}}$ and $D_{\mathrm{Primes}}$ are one edge read at two scales. Observed at finite working precision $P$, the
same law is the matching-digit count
$D(\lambda^2,N,P)=-\log_{10}(|\nu_1-\gamma_1|/|\gamma_1|)$,
\[
  D(\lambda^2,N,P)=\min\!\bigl(D_{\mathrm{Wprec}}(P),\,
  D_{\mathrm{nModes}}(N,\lambda^2),\,D_{\mathrm{Primes}}(\lambda^2)\bigr),
\]
the minimum of three budgets (precision, modes, and prime content), of which
precision is the instrument resolution, absent from the intrinsic form. In digit
space the minimum is forced, not fitted ($D$ is a negative logarithm of a sum of
independent errors, so the largest binds), which is why earlier single-expression
digit fits fail. We prove, under an explicit finite-cutoff positivity
hypothesis (unconditional in the Yoshida--Bombieri small-support range), that
convergence to $\gamma_1$ requires all three budgets to diverge: at a fixed
cutoff the accuracy is capped at the prime ceiling
$D_{\mathrm{Primes}}(\lambda^2)$, so no basis size or precision reaches the zero
and the cutoff must grow. This reduces convergence to the growth of the prime
ceiling, whose leading coefficient is derived, parameter-free, from the prolate
spheroidal eigenvalue asymptotic: $D_{\mathrm{Primes}}\sim(4\pi/\ln 10)\lambda^2$,
rigorously anchored from one side by a variational floor bound; the reverse
inequality, that the ceiling cannot be beaten at this rate, remains open. The
mode budget, the basis-size convergence rate, is the logarithm of a stretched-exponential
convergence $\exp(-aN^{q})$ whose exponent $q(\lambda^2)$ is the analyticity
class of the eigenvector, derived from the archimedean Gamma factor with a
prime correction from Mertens' theorem; a by-product is that the
eigenvalue-space Galerkin exponent measured by Groskin~\cite{Groskin2026} is a
finite-window slice of the same curve. We state each component with an explicit
rigor level, separating
what is proved versus what is derived and empirically validated up to
$\lambda^2=200$; the structural inputs are independent
of~\cite{CvS2025} (Appendix~\ref{app:selfcontained}). The results have been
communicated to co-authors of~\cite{CCM2025}.
\end{abstract}

\section{Introduction}\label{sec:intro}

The Connes--Consani--Moscovici (CCM) construction~\cite{CCM2025}, building on
the Connes--van Suijlekom truncated Weil form~\cite{CvS2025}, produces a family
of finite self-adjoint operators $D_{\log}^{(\lambda,N)}$ whose spectra
approximate the imaginary parts $\gamma_k$ of the nontrivial Riemann zeta
zeros to remarkable numerical accuracy: $55$ digits at the headline
configuration of~\cite{CCM2025}, which we extended to $999$ digits in an
independent reproduction~\cite{AndrewsA}. The construction has two integer-like
control parameters, a prime cutoff $\lambda^2$ (the primes $p\le\lambda^2$
enter the Weil form) and a basis size $N$ (the matrix is $(2N{+}1)\times(2N{+}1)$),
together with the working precision $P$ in decimal digits. The rate at
which the eigenvalue converges in the basis size $N$ is named the central open
problem in~\cite{CCM2025}; Groskin~\cite{Groskin2026}
measured a Galerkin convergence exponent empirically but disclaims deriving it,
and \'Sliwi\'nski~\cite{Sliwinski2026} proves an inverse-logarithmic lower bound
on the mean spectral error. We previously
tested two proposed rate laws against high-precision data and found them
unsupported~\cite{AndrewsB}.

The author's starting point is a belief the Hilbert--P\'olya conjecture
already expresses: that apparent randomness in the zeros is structure not yet
written down. The CCM construction gives that belief a computable form. What
remained open is the rate at which the computation converges, and
characterizing that rate explicitly is the subject of this paper.
\begin{definition}\label{def:D}
$D(\lambda^2,N,P) := -\log_{10}\bigl(|\nu_1-\gamma_1|/|\gamma_1|\bigr)$ is the
\emph{matching-digit count}: the number of correct leading decimal digits of the
first Riemann zero $\gamma_1$ reproduced by $\nu_1$, the first ordinate produced
by the CCM construction, computed at working precision $P$ via MPFR/GMP. The reference $\gamma_1$ is
computed to $2500$ significant digits by rigorous interval arithmetic
(Arb~\cite{Arb2017}), far exceeding the largest match reported here
($\approx1070$ digits at $\lambda^2=200$), so the reported counts are genuine and
not reference-limited; its leading digits agree with the standard tabulations
of~\cite{Odlyzko}.
\end{definition}

\begin{center}
\small
\begin{tabular}{@{}ll@{}}
\toprule
\multicolumn{2}{@{}l}{\textbf{Notation} (each object is defined where cited; collected here for reference)}\\
\midrule
$E=|\nu_1-\gamma_1|/|\gamma_1|$ & relative error of the first ordinate; $D=-\log_{10}E$ (Def.~\ref{def:D})\\
$\varepsilon_{\mathrm{modes}},\ \varepsilon_{\mathrm{primes}}$ & intrinsic error channels, eq.~\eqref{eq:channels}\\
$D_{\mathrm{Wprec}},\,D_{\mathrm{nModes}},\,D_{\mathrm{Primes}}$ & per-source digit budgets (Sections~\ref{sec:wprec}--\ref{sec:primes})\\
$\varepsilon^\infty(\lambda)$ & smallest eigenvalue of the continuum form at cutoff $\lambda$\\
$D_{\mathrm{eig}}=-\log_{10}\varepsilon^\infty$ & eigenvalue floor (Section~\ref{sec:primes})\\
$D_{\mathrm{edge}}=D_{\mathrm{Primes}}\cdot g$ & the single edge function (Section~\ref{ssec:twoforms})\\
\bottomrule
\end{tabular}
\end{center}

\noindent The same number is read as a \emph{ceiling} or a
\emph{floor} depending on the direction of view. $D_{\mathrm{Primes}}$ is a
ceiling on the achievable digit count and, equally, the depth of the accuracy
floor at the cutoff; both usages refer to the one object.

\noindent Our central result, the \emph{Andrews CCM Convergence Law}, has two
equivalent forms. Its intrinsic (spectral) form is the two-channel decomposition
\begin{equation}\label{eq:main-intrinsic}
  \boxed{\;E(N,\lambda^2)\le\varepsilon_{\mathrm{modes}}(N,\lambda^2)
         +\varepsilon_{\mathrm{primes}}(\lambda^2)\;}
\end{equation}
with the channels defined in Section~\ref{sec:skeleton},
\eqref{eq:channels} (an equality up to sign cancellation, confined to
crossover seams; Prop.~\ref{prop:min}). The two channels are the finite-$N$ transition and the extreme tail of a single
prolate-type spectral edge (Section~\ref{ssec:twoforms}); convergence
$\nu_1\to\gamma_1$ is exactly $E\to0$. Its finite-precision measurement is the
matching-digit count
\begin{equation}\label{eq:main}
  \boxed{\;D(\lambda^2,N,P)=\min\!\bigl(D_{\mathrm{Wprec}}(P),\,
  D_{\mathrm{nModes}}(N,\lambda^2),\,D_{\mathrm{Primes}}(\lambda^2)\bigr)\;}
\end{equation}
in which $D_{\mathrm{Wprec}}$ is the instrument resolution alone. Since $D$ is the
relative error on a logarithmic scale, the law is a quantitative convergence
statement: $D\to\infty$ is precisely $\nu_1\to\gamma_1$, and the three budgets
are the independent mechanisms that drive (or stall) that convergence
(Prop.~\ref{prop:limit}). Five points
distinguish it from prior empirical fits:
\begin{enumerate}
\item \textbf{The minimum is forced, not fitted} (Section~\ref{sec:skeleton}).
  In digit space $D$ is a negative logarithm of an error bounded by a sum of
  operationally independent channels, so the largest binds and $D$ is at least
  the minimum of the three per-source budgets, unconditionally, and equals it up
  to an $O(1)$ term absent cancellation (Prop.~\ref{prop:min}). This is the
  structural reason earlier single-formula digit fits failed. The precision budget is measurement, not
  mathematics: removing it leaves the single intrinsic
  edge~\eqref{eq:main-intrinsic}, whose two channels (modes, primes) are the only
  mathematical content.
\item \textbf{Each budget is analyzed from the construction}, not guessed. The
  precision budget follows from elementary round-off analysis; the prime
  ceiling from the prolate eigenvalue asymptotic; and the mode budget from the
  analyticity of the Weil-form eigenvector.
\item \textbf{Every claim carries an explicit rigor level} (Table~\ref{tab:tiers}),
  separating the rigorously proved parts from those derived and validated
  against high-precision computation. We do not label the whole law a
  theorem; we state precisely which pieces are.
\item \textbf{It yields a convergence criterion} (Prop.~\ref{prop:limit}): the
  prime cutoff is an unbeatable ceiling. At a fixed $\lambda^2$ no basis size
  or precision reaches the zero, so convergence to $\gamma_1$ forces
  $\lambda^2\to\infty$, which turns the qualitative convergence
  of~\cite{CCM2025} into an explicit, conditional rate.
\item \textbf{No prior unified law, to our knowledge.} Earlier attempts are
  single closed-form fits of the digit count, and they fail: we tested two
  proposed rate laws against high-precision data and found them
  unsupported~\cite{AndrewsB}. Groskin~\cite{Groskin2026} likewise reports
  that no single smooth-convergence law fits his data and declines to assert a
  rate, and \'Sliwi\'nski~\cite{Sliwinski2026} establishes only a lower bound
  on the mean error. The unified characterization given here, the intrinsic
  two-channel edge~\eqref{eq:main-intrinsic} and its finite-precision
  measurement~\eqref{eq:main}, is, to the best of the author's knowledge, the
  first working closed-form convergence law for the construction. Its unity is
  the point: one spectral edge (Section~\ref{sec:equilibrium}) underlies both the
  mode ramp and the prime ceiling, which is why a single expression can hold
  where earlier per-fits could not.
\end{enumerate}

The deepest new content is the mode budget $D_{\mathrm{nModes}}$
(Section~\ref{sec:nmodes}): we show the matching-digit ramp is the logarithm of
a stretched-exponential convergence, identify its exponent with the analyticity
class of the eigenvector, and derive that exponent end to end. A by-product is
that the eigenvalue-space Galerkin exponent measured by Groskin~\cite{Groskin2026}
is a finite-window slice of the same curve. That the mode ramp and the prime
ceiling are not two mechanisms but one spectral edge at two scales, the lower
edge of an equilibrium measure whose bulk is the Riemann--von Mangoldt density,
is the subject of Section~\ref{sec:equilibrium}.

\section{The construction and the object $D$}\label{sec:background}

We recall the construction of~\cite{CCM2025,CvS2025}. Let $\lambda>1$ and
$L=2\ln\lambda=\ln\lambda^2$. On $\mathcal H_\lambda=L^2([\lambda^{-1},\lambda],
du/u)$ the functions $V_n(u)=L^{-1/2}e^{2\pi i n\ln(\lambda u)/L}$ form an
orthonormal basis. The Weil quadratic form $QW_\lambda$, restricted to
$\mathrm{span}\{V_n:|n|\le N\}$, is a real symmetric $(2N{+}1)\times(2N{+}1)$
matrix whose archimedean part carries the local factor at infinity and whose
arithmetic part is a finite sum over prime powers $p^j\le\lambda^2$. Its
smallest eigenvalue $\varepsilon_N$ has an even eigenvector $\xi$, and the
positive zeros of the associated Fourier--Mellin function $\widehat\xi$
approximate the ordinates $\gamma_k$~\cite{CCM2025}.

Two facts about $\widehat\xi$ are used below: $\widehat\xi$ is entire, and (in
the critical case) all its zeros are real and coincide with the spectrum. Both
are proven \emph{self-contained} in~\cite{CCM2025}: the closed form of the
transform is \cite[Prop.~5.9]{CCM2025} (rederived from scratch in
Appendix~\ref{app:selfcontained}), and entirety/reality is
\cite[Thm.~5.10]{CCM2025}, whose proof chain (Lemmas~5.2 and~5.4 there: the
matrix structure $\tau_{ij}=(b_i-b_j)/(i-j)$, an elementary commutator
identity, and self-adjointness in the induced inner product) is complete
within~\cite{CCM2025}. The Carath\'eodory--Fej\'er--type reality mechanism
\emph{originates} in the Connes--van Suijlekom paper~\cite{CvS2025}, which we
cite as the historical source; no statement used in this paper depends on it
(Appendix~\ref{app:selfcontained}). An independent classical foundation for the
truncated Weil form predates both: Weil's positivity criterion, Yoshida's
unconditional positivity theorem for test functions of sufficiently small
support~\cite{Yoshida1992}, Bombieri's study of the truncated quadratic form
and the eigenvalues of its finite sections~\cite{Bombieri2000}, and Suzuki's
screw-function framework~\cite{SuzukiJLMS,SuzukiHilbert,Suzuki2026}, which
unifies all of these by means of continuous functions.

The eigenvector $\xi$ is even and, in the regime we
study, the smallest eigenvalue is simple, the even-symmetry verified
numerically across $38$ configurations spanning $\lambda^2=13$--$1000$
in~\cite{AndrewsA}.

The object of study, $D(\lambda^2,N,P)$, is a digit count: $D=-\log_{10}
(\text{relative error})$. The relative error itself,
$E=|\nu_1-\gamma_1|/|\gamma_1|$, is the intrinsic object of the
law~\eqref{eq:main-intrinsic}; $D$ is its measurement in digits, and the
arithmetic of that measurement is what makes the minimum structure
of~\eqref{eq:main} inevitable.

\section{The three-budget skeleton}\label{sec:skeleton}

The computed approximation $\nu_1$ carries error from three sources:
finite working precision (round-off), finite basis size (mode truncation), and
finite prime cutoff (the intrinsic ceiling at $N,P\to\infty$). We now
\emph{define} the channels, rather than posit their additivity. Write
$\nu_1(N,\lambda^2)$ for the first ordinate of the exact (infinite-precision)
finite-$N$ operator and $\nu_1(\infty,\lambda^2)$ for its $N\to\infty$ limit at
fixed cutoff, and set
\begin{equation}\label{eq:channels}
  \varepsilon_{\mathrm{modes}}(N,\lambda^2)
   :=\frac{|\nu_1(N,\lambda^2)-\nu_1(\infty,\lambda^2)|}{|\gamma_1|},
  \qquad
  \varepsilon_{\mathrm{primes}}(\lambda^2)
   :=\frac{|\nu_1(\infty,\lambda^2)-\gamma_1|}{|\gamma_1|},
\end{equation}
the finite-$N$ displacement at fixed cutoff and the cutoff displacement at
infinite basis size, with $\varepsilon_{\mathrm{round}}(P)$ the round-off
perturbation of the computed $\nu_1$ relative to the exact finite-$N$ value.
The triangle inequality then gives, for the computed relative error,
\[
  E \;\le\; \varepsilon_{\mathrm{round}}(P)
  +\varepsilon_{\mathrm{modes}}(N,\lambda^2)+\varepsilon_{\mathrm{primes}}(\lambda^2),
\]
an inequality, not an identity: the signed displacements can in principle
cancel. The channels are \emph{independent} in a precise operational (not
probabilistic or structural) sense: each answers to its own parameter
(round-off to $P$, truncation to $N$, prime content to $\lambda^2$), so any
one can be driven below the others at will. (Structurally they are coupled:
$\varepsilon_{\mathrm{modes}}$ depends on $\lambda^2$ through both the exponent
$q$ and the Shannon coordinate of Section~\ref{sec:nmodes}, and below the
precision floor round-off corrupts the eigenvector nonlinearly,
Section~\ref{ssec:evenness}; the operational sense is all
Prop.~\ref{prop:min} uses.)

\begin{proposition}[The minimum is forced]\label{prop:min}
Let $\varepsilon_1,\varepsilon_2,\varepsilon_3\ge0$ and let $E\ge0$ satisfy
$E\le\Sigma:=\sum_i\varepsilon_i$. Then, unconditionally,
\[
  D:=-\log_{10}E\;\ge\;-\log_{10}\Sigma
  \;\ge\;\min_i\bigl(-\log_{10}\varepsilon_i\bigr)-\log_{10}3 ,
\]
since $\Sigma\le 3\max_i\varepsilon_i$. If in addition no cancellation occurs
--- quantitatively, $E\ge\theta\max_i\varepsilon_i$ for some fixed
$\theta\in(0,1]$ --- then also
$D\le\min_i(-\log_{10}\varepsilon_i)+\log_{10}(1/\theta)$, whence
$D=\min(D_{\mathrm{Wprec}},D_{\mathrm{nModes}},D_{\mathrm{Primes}})+O(1)$.
\end{proposition}

The lower bound, that the computed digit count is at least the binding budget
minus half a digit, is a rigorous consequence of the triangle inequality
alone. The matching upper bound holds \emph{absent cancellation} between the
signed displacements. Such cancellation is confined to the narrow crossover
seams where two budgets are comparable (away from a seam a single channel
dominates $E$ outright and $\theta$ is near $1$), and no seam in the
$5{,}000$-plus experiments of this program shows $D$ exceeding the binding
minimum by more than the stated $O(1)$; with that caveat recorded once, we
write the law as an equality throughout. The $O(1)$ term is at most
$\log_{10}3<0.48$ of a digit and lives only at the seams, so the law's content
is that the \emph{binding mechanism} changes across the parameter space
(round-off, mode truncation, then prime content), not that $D$ is literally
non-smooth (it is smooth: a negative logarithm of a near-sum).

This is why, in digit space, $D$ is a minimum rather than a sum or product, and
why global single-expression fits of the digit count fail in practice: a single
elementary term cannot simultaneously hold a hard precision wall, a prime
ceiling, and a $\lambda^2$-coupled mode ramp. In the intrinsic form the error
is, by contrast, the two-channel decomposition~\eqref{eq:main-intrinsic}; the
digit-space minimum
is what the negative logarithm of that near-sum becomes, the largest term
binding. The junctions differ in character.
The precision junction is a hard wall, since round-off is a clamp, so its corner
is sharp. The mode$\to$prime junction, by contrast, is a wide, smooth saturation
(the $\tanh$ of Section~\ref{sec:nmodes}): $D_{\mathrm{nModes}}$ is the
mode-limited digit count, which rises along its pre-saturation ramp and is capped
by the ceiling, $D_{\mathrm{nModes}}\le D_{\mathrm{Primes}}$, with equality as
$N\to\infty$.

\paragraph{The starvation protocol.}
Each budget is isolated experimentally by \emph{starvation}: to read $D_i$
directly, set the other two parameters large enough that their budgets exceed
$D_i$, so the channel of interest dominates $E$ outright and the cancellation
caveat of Prop.~\ref{prop:min} is vacuous. This protocol is the
experimental mirror of the minimum structure. It is what makes the digit form
falsifiable rather than descriptive, since each budget's functional form is
measured in a regime where it alone binds. We use it throughout
Sections~\ref{sec:wprec}--\ref{sec:primes}.

\subsection{The two forms of the law}\label{ssec:twoforms}

Of the three budgets, only two (mode truncation and prime cutoff) are
properties of the finite operator; round-off is a property of the computation.
Separating them gives the two forms stated in Section~\ref{sec:intro}.

\smallskip
\noindent\emph{Intrinsic (spectral) form.} The mathematical content is the
two-channel decomposition~\eqref{eq:main-intrinsic}, with
$\varepsilon_{\mathrm{modes}}\sim
\exp(-aN^{q})\to0$ as $N\to\infty$ (Section~\ref{sec:nmodes}) and
$\varepsilon_{\mathrm{primes}}\to0$ as $\lambda^2\to\infty$
(Section~\ref{sec:primes}). It makes no reference to precision, and
$\nu_1\to\gamma_1$ is exactly $E\to0$. Vanishing of both channels forces
$E\to0$ unconditionally; conversely, sustained convergence $E\to0$ forces both
channels to vanish, since a bounded cutoff pins $\varepsilon_{\mathrm{primes}}$
away from zero (Prop.~\ref{prop:limit}) and the fixed-$\lambda^2$ mode
displacement, vanishing as $N\to\infty$, cannot cancel it in the limit ---
accidental cancellation can occur only at isolated parameter values, not along
a convergent path.

\smallskip
\noindent\emph{Observed (digit) form.} A computation at working precision $P$
resolves $E$ only down to its round-off floor, so the measured matching-digit
count is
\begin{equation}\label{eq:observed}
  D(\lambda^2,N,P)=\min\!\bigl(D_{\mathrm{Wprec}}(P),\,-\log_{10}E(N,\lambda^2)\bigr)
   =\min\!\bigl(D_{\mathrm{Wprec}},D_{\mathrm{nModes}},D_{\mathrm{Primes}}\bigr)+O(1),
\end{equation}
the second equality by Prop.~\ref{prop:min}. Here $D_{\mathrm{Wprec}}(P)$
is the instrument resolution: it is no property of $\nu_1$, $\gamma_1$, or the
operator, and as $P\to\infty$ it disappears, returning the intrinsic form in
digit units. Thus the boxed law~\eqref{eq:main} is the finite-precision
\emph{measurement} of the intrinsic law~\eqref{eq:main-intrinsic}.

\smallskip
\noindent\emph{One edge, not two budgets.} The mode and prime channels are the
two ends of a single spectral edge. Dividing out the ceiling
(Section~\ref{sec:nmodes}) writes $D_{\mathrm{nModes}}=D_{\mathrm{Primes}}\cdot g$
with $g=\tanh\!\bigl((c\,N/(\lambda^2\ln\lambda^2))^{q}\bigr)\uparrow1$ as
$N\to\infty$, so $D_{\mathrm{Primes}}=\lim_{N\to\infty}D_{\mathrm{nModes}}$ is the
$N\to\infty$ ceiling of the mode ramp, not a separate term; equivalently the
digit law is a single edge function $D_{\mathrm{edge}}(N,\lambda^2)=
D_{\mathrm{Primes}}\cdot g$ capped by precision,
$D=\min(D_{\mathrm{Wprec}},D_{\mathrm{edge}})+O(1)$. Correspondingly the intrinsic
error~\eqref{eq:main-intrinsic} is one edge: $\varepsilon_{\mathrm{modes}}$ is its
finite-$N$ transition (the Landau--Widom plunge, Section~\ref{sec:nmodes}) and
$\varepsilon_{\mathrm{primes}}$ is its extreme tail (the prolate asymptotic,
Section~\ref{sec:primes}). That these are literally the transition and the tail
of one band edge, the lower edge of the archimedean equilibrium measure,
is established in Section~\ref{sec:equilibrium}.

\section{The precision budget $D_{\mathrm{Wprec}}$}\label{sec:wprec}

Finite working precision $P$ (decimal digits) introduces relative round-off
$\sim 10^{-P}$, amplified by the eigenproblem's condition number $\kappa$ (set
by the gap of $\varepsilon_N$ to the rest of the spectrum and independent of
$P$). Hence
\begin{equation}\label{eq:wprec}
  D_{\mathrm{Wprec}}(P)=P-\log_{10}\kappa = P+O(1),
\end{equation}
with leading coefficient exactly $1$. The $O(1)$ term is implementation
dependent: the toolkit used here allocates $\lceil P\cdot\ln10/\ln2\rceil+16$
working bits, the $16$ guard bits contributing a small positive offset. Isolating
this budget (large $\lambda^2$ and $N$) and sweeping $P\in\{100,200,300,400,700,
1000\}$ gives a measured slope $\partial D/\partial P=1.000\pm0.001$. The exact
offset is not a property of the construction and we do not assign it physical
meaning. \emph{Status: T1; the coefficient $1$ follows from elementary numerical
analysis.}

\subsection{Symmetry, positivity, and the precision floor}\label{ssec:evenness}

A question the precision floor settles, and one easy to answer wrongly
without it, is whether the even symmetry of the ground eigenvector, and the
positivity of $\varepsilon_N$, are genuine features of the operator or merely of
the regime above the floor. The CCM construction~\cite{CCM2025} posits a simple
smallest eigenvalue with an \emph{even} eigenvector (its Step~1) and enforces
this with a forced-even projection. The reproduction of~\cite{AndrewsA} tested
that projection directly: across $38$ configurations ($\lambda^2=13$--$1000$,
$N=10$--$800$, spanning both sides of the precision floor) with the projection
\emph{disabled}, the natural smallest eigenvector converged to the very vector
the projection would have produced, to \emph{bit-identical} eigenvalues,
at every configuration above the floor. The forced-even projection is thus
empirically a no-op above the floor: even symmetry is a property the operator
exhibits unbidden, not an artifact of the projection or of the above-floor
regime.

The qualifier ``above the floor'' is essential, and it is precisely where the
budget of this section bites. When the target eigenvalue lies below
$D_{\mathrm{Wprec}}$, smaller than the round-off that the working precision
can resolve, the computed eigenvector is dominated by numerical noise: it can
present spurious mixed symmetry, and the smallest computed eigenvalue can carry
a spurious sign. These are artifacts of precision starvation, not properties of
the operator; raising $P$ past the floor dissolves them, and the natural
eigenvector is again even, into the deep near-zero spectrum. (This is also the
regime in which the round-off channel of Section~\ref{sec:skeleton} stops
acting as an additive perturbation of $\nu_1$: below the floor it corrupts the
eigenvector nonlinearly, so the decomposition~\eqref{eq:channels}, like the
symmetry and positivity statements of this subsection, is an above-floor
statement.)

This even, positive ground state was documented in our reproduction~\cite{AndrewsA}.
In subsequent correspondence, and in a revision of his own paper~\cite{Groskin2026},
Groskin independently corroborated it by a different route: the sign changes his
separate quadrature implementation had shown among the deepest eigenvalues at a
finite archimedean cutoff $T$ were identified as a finite-$T$ artifact that
vanishes as $T$ is increased, consistent with the naturally even, positive ground
state of~\cite{AndrewsA}. Working precision $P$ and archimedean cutoff $T$ are
therefore independent controls: a deep-spectrum value is trustworthy only when it
is stable across two settings of \emph{each}, not when it merely holds steady in
one. The even, positive ground state persists into the deep near-zero spectrum;
apparent breakdowns are finite-precision or finite-cutoff artifacts, not features
of the construction.

\section{The mode budget $D_{\mathrm{nModes}}$}\label{sec:nmodes}

This is the convergence rate in $N$, the open problem of~\cite{CCM2025}. In
the intrinsic form it is the mode channel $\varepsilon_{\mathrm{modes}}$; as a
digit count it is the rising transition of the equilibrium-measure edge
(Section~\ref{sec:equilibrium}), climbing toward the prime ceiling.
Dividing out that ceiling (Section~\ref{sec:primes}), write the ramp ratio
$g=D_{\mathrm{nModes}}/D_{\mathrm{Primes}}\in[0,1]$. We claim
\begin{equation}\label{eq:nmodes}
  D_{\mathrm{nModes}}(N,\lambda^2)=D_{\mathrm{Primes}}(\lambda^2)\cdot
  \tanh\!\Bigl(\bigl(c\,N/(\lambda^2\ln\lambda^2)\bigr)^{q(\lambda^2)}\Bigr),
  \qquad q(\lambda^2)=\frac{2}{2\gamma+\ln\ln\lambda^2}.
\end{equation}
We build this in four steps: the stretched-exponential meaning of the ramp
(\ref{ssec:stretch}), the archimedean leading rate (\ref{ssec:arch}), the prime
correction giving $q$ (\ref{ssec:prime}), and the Shannon coordinate
(\ref{ssec:coord}).

\subsection{The ramp is a stretched exponential}\label{ssec:stretch}

\begin{proposition}[Stretched-exponential convergence and rate unification]\label{prop:stretched}
Suppose the pre-saturation mode budget satisfies $D_{\mathrm{nModes}}=(a/\ln10)\,
N^{q}(1+o(1))$ for some $a=a(\lambda^2)>0$ and exponent $q=q(\lambda^2)$, and that
the Weil-form ground eigenvalue is simple with even eigenvector. Then:
\begin{enumerate}
\item[(i)] \emph{(Stretched-exponential.)} The eigenvalue/zero error obeys
  $|\nu_1-\gamma_1|/|\gamma_1|=\exp(-aN^{q})(1+o(1))$; equivalently $q$ is the
  analyticity class of the eigenvector ($q=1$ exponential / strip-analytic,
  $0<q<1$ sub-exponential but faster than any polynomial, $q\to0$ finite Sobolev)
  --- on the classical scale, $q=1/\mu$ with $\mu$ the Gevrey index of the
  eigenvector \cite{Rodino1993}.
\item[(ii)] \emph{(Rate unification.)} Any local power-law fit
  $\mathrm{err}\sim C\,N^{-2s}$ over a window about $N$ returns the \emph{growing}
  local exponent $2s_{\mathrm{loc}}(N)=-\,d\ln\mathrm{err}/d\ln N=a\,q\,N^{q}$.
  Hence the eigenvalue-space Galerkin exponent $s(c)$ of~\cite{Groskin2026} is the
  local logarithmic slope of this same stretched-exponential --- not a fixed
  Sobolev index --- and the digit-space ramp and the eigenvalue-space rate are one
  object in two coordinates. The factor $2$ is the eigenvalue-approximation
  identity $|\varepsilon-\varepsilon_N|=O(\|\xi-\xi_N\|^2)$ for a self-adjoint
  operator with simple ground state~\cite{BabuskaOsborn1991}.
\end{enumerate}
\end{proposition}
\begin{proof}
(i) $D=-\log_{10}(\mathrm{err})\Rightarrow\mathrm{err}=10^{-D}=\exp(-aN^q)$.
(ii) $2s_{\mathrm{loc}}=-\,d\ln\mathrm{err}/d\ln N=d(aN^q)/d\ln N=aqN^q$; the
squaring is the Rayleigh-quotient second-order identity at a simple
eigenvalue~\cite{BabuskaOsborn1991}. A degenerate ground state would give only
first-order dependence and no factor $2$.
\end{proof}

Thus $q$ is not a fitting knob but the \emph{analyticity class} of the
eigenvector: $q=1$ is pure exponential decay (analytic in a strip), $0<q<1$ is
sub-exponential but faster than any polynomial, and $q\to0$ is finite Sobolev
($N^{-2s}$). In standard terminology this scale is the \emph{Gevrey scale}: the
Gevrey class $G^{\mu}$, $\mu\ge1$, is characterized exactly by Fourier-coefficient
decay $\exp(-c\,|k|^{1/\mu})$, reducing to strip-analytic functions at $\mu=1$
\cite{Rodino1993}, so the statement $|\xi_k|\sim\exp(-c\,k^{q})$ is the
statement $\xi\in G^{1/q}$ (and no better). The open ``variance-to-exponent
map'' of Section~\ref{ssec:prime} is thereby a precisely posed regularity
lemma on a classical scale: \emph{prove that the ground eigenvector lies in the
Gevrey class $G^{1/q(\lambda^2)}$ sharply, with
$1/q=\gamma+\tfrac12\ln\ln\lambda^2$}. The measured exponents lie just below $1$
and decrease with
$\lambda^2$ (Table~\ref{tab:q}), the signature of an eigenvector that is
strip-analytic in its archimedean part and roughened by the prime content.

The force of part~(i) is that it converts the open problem of \emph{measuring}
a convergence rate into the analytic question of \emph{identifying} the
eigenvector's regularity: the single exponent $q$ fixes the entire pre-saturation
curve. It also explains why fixed power-law (Sobolev) fits are structurally the
wrong model (they are the $q\to0$ limit, whereas the operator sits near $q=1$),
and hence why such fits drift and resist extrapolation
(Section~\ref{sec:discussion}).

Part~(ii) identifies the Galerkin exponent $s(c)$ measured by
Groskin~\cite{Groskin2026} in eigenvalue space as the local logarithmic slope
of the present stretched-exponential, sampled at his window; the data-level
reconciliation this affords is taken up in Section~\ref{sec:discussion}. The squaring
that links them needs the ground eigenvalue to be \emph{simple}, which is not
assumed but established as the naturally-even, isolated ground state verified
across $38$ configurations in the reproduction~\cite{AndrewsA}; were the state
degenerate, the eigenvalue error would be only first order in the eigenvector
error and the factor $2$, equivalently the $2s$ in Groskin's rate, would not
appear. \emph{Status: T1 (given the pre-saturation ramp form as hypothesis).}

\subsection{The archimedean leading rate $\delta=\pi/4$}\label{ssec:arch}

The first positive zero $\nu_1$ of the Fourier--Mellin function $\widehat\xi(z)$
converges to $\gamma_1$ at a rate set by the analyticity strip of the
\emph{limiting} Fourier--Mellin function, governed by the archimedean factor of
the Weil form. That factor
is the completed real Gamma factor $\Gamma_\mathbb{R}(s)=\pi^{-s/2}\Gamma(s/2)$:
its
logarithmic derivative $\psi(\tfrac14+\tfrac{iz}{2})\sim\ln(z/2)$ is precisely
the $\ln|z|$ multiplier identified by Suzuki~\cite{Suzuki2026}, and the factor
itself decays as $|\Gamma(\tfrac14+\tfrac{iz}{2})|\sim\exp(-\pi|z|/4)$ by
Stirling. Hence the strip half-width is
\begin{equation}\label{eq:delta}
  \delta=\pi/4 .
\end{equation}

This value has a clean structural origin, one that also shows why the scales
$\tfrac12$ and $\pi/4$ both attach to the same archimedean factor. On the
critical line the archimedean log-derivative has the large-$z$ expansion
\begin{equation}\label{eq:digamma-split}
  \psi\!\left(\tfrac14+\tfrac{iz}{2}\right)=\log\frac{z}{2}
  +i\,\frac{\pi}{2}+o(1),\qquad z\to+\infty,
\end{equation}
since the argument $\tfrac14+\tfrac{iz}{2}$ runs in the imaginary
(critical-line) direction, so $\arg\bigl(\tfrac14+\tfrac{iz}{2}\bigr)\to\pi/2$.
Its real and imaginary parts carry the two scales:
\begin{itemize}
\item the \emph{real} part $\operatorname{Re}\psi(\tfrac14+\tfrac{iz}{2})\to
  \log(z/2)$ is the symbol, Suzuki's $\log|z|$ multiplier~\cite{Suzuki2026},
  whose analyticity strip has half-width $\tfrac12$ (the digamma poles at
  $z=\pm i/2$);
\item the \emph{imaginary} part $\operatorname{Im}\psi(\tfrac14+\tfrac{iz}{2})\to
  \tfrac{\pi}{2}$ sets the envelope, through
  \[
    \frac{d}{dz}\log\bigl|\Gamma_\mathbb{R}(\tfrac12+iz)\bigr|
    =-\tfrac12\operatorname{Im}\psi\!\left(\tfrac14+\tfrac{iz}{2}\right)
    \longrightarrow-\tfrac12\cdot\tfrac{\pi}{2}=-\tfrac{\pi}{4}.
  \]
\end{itemize}
The strip half-width and the envelope rate are therefore the real and imaginary
parts of the \emph{one} analytic function $\log\Gamma_\mathbb{R}(\tfrac12+iz)$,
a Hilbert (Kramers--Kronig) conjugate pair since $\Gamma_\mathbb{R}$ is zero-free
(minimum phase): its log-magnitude and phase determine each other. In one
identity,
\begin{equation}\label{eq:pi4}
  \delta=\frac{\pi}{4}=\frac12\cdot\frac{\pi}{2}=\tfrac12\arg(i),
\end{equation}
half the right angle of the critical-line direction. The two scales $\tfrac12$
and $\pi/4$ are thus not independent constants but conjugate parts of the single
factor $\Gamma_\mathbb{R}$: the strip half-width $\tfrac12$ and its harmonic
conjugate $\pi/4$.

Two points delimit what this establishes. It fixes the \emph{value}
$\delta=\pi/4$ as a property of the limiting $\Gamma_\mathbb{R}$ factor: $\delta$
is the half-width of the strip carrying the zeros of the limiting Fourier--Mellin
function, and governs the zero convergence $\nu_1\to\gamma_1$ (with the squaring
of Prop.~\ref{prop:stretched}(ii)); it is not the decay rate of the finite
eigenvector's Galerkin coefficients, a distinct band-limited, prime-modulated
quantity ($\approx0.71$ at $\lambda^2=100$), and the finite $\widehat\xi$, of
exponential type $L/2$ via its $\sin(zL/2)$ factor, is not literally the completed
$\xi$. And it is structural motivation rather than a full derivation: obtaining
$\pi/4$ from the symbol alone would require its conjugate part, which appears only
through the analytic completion $\Gamma_\mathbb{R}$, that is, through the finite
ground-state transform inheriting that factor, the realization step that remains
open.

\begin{lemma}[Lattice sampling of the coefficients]\label{lem:sampling}
The Fourier--Mellin transform of the $N$-mode eigenvector has the closed
form $\widehat\xi_N(z)=2L^{-1/2}\sin(zL/2)\sum_{|j|\le N}\xi_j/(z-2\pi j/L)$
(\cite[Prop.~5.9]{CCM2025}; rederived from scratch in
Appendix~\ref{app:selfcontained}). On the frequency lattice $z_k=2\pi k/L$,
\[
  \widehat\xi_N(2\pi k/L)=(-1)^k\sqrt{L}\,\xi_k\ \ (|k|\le N),\qquad
  \widehat\xi_N(2\pi k/L)=0\ \ (|k|>N),
\]
so $|\xi_k|=L^{-1/2}\,\bigl|\widehat\xi_N(2\pi k/L)\bigr|$, with
$L=\ln\lambda^2$ the period of the basis $\{V_n\}$.
\end{lemma}
\begin{proof}
At $z_k=2\pi k/L$ the factor $\sin(zL/2)=(-1)^k\sin\!\bigl((z-z_k)L/2\bigr)$
has a simple zero, so every $j\ne k$ term vanishes and the $j=k$ pole is
removable:
$\lim_{z\to z_k}2L^{-1/2}(-1)^k\sin\!\bigl((z-z_k)L/2\bigr)/(z-z_k)
=(-1)^k\sqrt{L}$. For $|k|>N$ there is no $j=k$ term and the common factor
$\sin(z_kL/2)=0$.
\end{proof}

By Lemma~\ref{lem:sampling} the coefficients are the lattice samples of
$\widehat\xi_N$ at spacing $2\pi/L$ with $L=\ln\lambda^2$, so the strip
half-width~\eqref{eq:delta} gives a coefficient tail $\exp(-\delta(2\pi/L)N)$;
with the squaring of Prop.~\ref{prop:stretched}(ii) the leading digit slope is
\begin{equation}\label{eq:slope}
  \frac{\partial D_{\mathrm{nModes}}}{\partial N}
  =\frac{2\delta\,(2\pi/L)}{\ln10}=\frac{\pi^2}{\ln10\;\ln\lambda^2}
  \approx\frac{4.286}{\ln\lambda^2},
\end{equation}
a parameter-free prediction. At $\lambda^2=7$, where few primes enter and
$q\approx1$, \eqref{eq:slope} gives $2.203$ against a measured small-$N$ slope
of $2.264$, agreement to $3\%$ with no free parameters. At larger $\lambda^2$
the measured early slope runs above~\eqref{eq:slope} exactly as the $q<1$
concavity (below) requires.

What \emph{is} rigorous is the transfer from strip convergence to zero
displacement, which we now record; it confines the entire open content of
$\delta=\pi/4$ to a coefficient-decay estimate.

\begin{lemma}[Zero displacement from strip convergence]\label{lem:rouche}
Let $F$ be holomorphic on a neighborhood of the closed disc
$\overline{D}(\gamma_1,r)$ with a simple zero at $\gamma_1$, no other zero in the
disc, and $|F(z)|\ge\tfrac12|F'(\gamma_1)|\,|z-\gamma_1|$ on the disc (true for
$r$ small enough, by continuity of $F'$). Let $G$ be holomorphic on the disc
with $\eta:=\sup_{\overline{D}(\gamma_1,r)}|G-F|$. If
$\rho:=2\eta/|F'(\gamma_1)|\le r$, then $G$ has exactly one zero $\nu$ in
$D(\gamma_1,\rho)$, and
\[
  |\nu-\gamma_1|\;\le\;\frac{2\,\eta}{|F'(\gamma_1)|}.
\]
\end{lemma}
\begin{proof}
On the circle $|z-\gamma_1|=\rho$ one has
$|F(z)|\ge\tfrac12|F'(\gamma_1)|\rho=\eta\ge|G(z)-F(z)|$, with strict inequality
after shrinking $\rho$ infinitesimally if equality holds; Rouch\'e's theorem then
gives $G$ the same number of zeros as $F$ in the disc, namely one.
\end{proof}

Applied with $F=\widehat\xi_\infty$ (the fixed-$\lambda^2$ limit transform,
whose zeros are real and simple in the critical case, the Hermite--Biehler
structure underlying the de~Branges-space treatment of the Weil
form~\cite{SuzukiHilbert,SuzukiJLMS}) and $G=\widehat\xi_N$, the lemma converts
uniform convergence on a neighborhood of $\gamma_1$ inside the strip into zero
displacement with the explicit constant $2/|\widehat\xi_\infty'(\gamma_1)|$.

A second transfer can be made rigorous as well: from the coefficient tail to
the transform. The subtlety is that the finite-section eigenvector is not the
truncation of the limit eigenvector, so the transform difference carries a
retained-coefficient error in addition to the discarded tail; both pieces are
controlled by the tail alone.

\begin{lemma}[Tail-to-transform transfer]\label{lem:tail}
Fix $\lambda^2$, assume the ground state of the continuum form $QW_\lambda$ is
simple with spectral gap $g(\lambda^2)>0$, and suppose the limit eigenvector
satisfies $|\xi^{(\infty)}_j|\le C\exp(-c\,j^{q})$ for some $c>0$,
$0<q\le1$. Let $r>0$ be small enough that
$\overline{D}(\gamma_1,r)$ has positive distance $d(\lambda^2,r)$ from the
frequency lattice $\{2\pi j/L\}$. Normalize the eigenvectors and fix phases by
$\langle\xi^{(N)},\xi^{(\infty)}\rangle\ge0$. Then
\[
  \sup_{|z-\gamma_1|\le r}\bigl|\widehat\xi_N(z)-\widehat\xi_\infty(z)\bigr|
  \;\le\;\exp\!\bigl(-c\,N^{q}(1+o(1))\bigr),\qquad N\to\infty .
\]
\end{lemma}
\begin{proof}
Write $P_N$ for the projection onto $\mathrm{span}\{V_j:|j|\le N\}$ and
$Q_N=I-P_N$. With $\ell^1$ coefficients the closed form of
Lemma~\ref{lem:sampling} extends to the limit: the series
$\widehat\xi_\infty(z)=2L^{-1/2}\sin(zL/2)\sum_j\xi_j/(z-2\pi j/L)$ converges
locally uniformly (lattice poles removable as before) and defines an entire
function. Split
$\widehat\xi_N-\widehat\xi_\infty=\widehat{T}[\xi^{(N)}-P_N\xi^{(\infty)}]
-\widehat{T}[Q_N\xi^{(\infty)}]$, with $\widehat T$ the linear
coefficients-to-function map. \emph{Tail piece:} on the disc
$|\sin(zL/2)|\le\cosh(rL/2)$ and $|z-2\pi j/L|\ge\pi|j|/L$ for $|j|$ large,
so $|\widehat T[Q_N\xi^{(\infty)}]|\le
C_1\sum_{|j|>N}e^{-cj^{q}}/|j|\le C_2N^{-q}e^{-cN^{q}}(1+o(1))$, by comparison
with $\int_N^\infty e^{-ct^{q}}\,dt$. \emph{Head piece:} the Galerkin residual
collapses to the tail,
\[
  \bigl(QW^N_\lambda-\varepsilon^\infty\bigr)\bigl(P_N\xi^{(\infty)}\bigr)
  =P_NQW_\lambda\xi^{(\infty)}-P_NQW_\lambda Q_N\xi^{(\infty)}
   -\varepsilon^\infty P_N\xi^{(\infty)}
  =-P_NQW_\lambda Q_N\xi^{(\infty)},
\]
using $QW_\lambda\xi^{(\infty)}=\varepsilon^\infty\xi^{(\infty)}$. Row bounds
for the form (diagonal $O(\log j)$ from the archimedean symbol, off-diagonal
rows uniformly $\ell^2$, prime part bounded) give
$\|QW_\lambda Q_N\xi^{(\infty)}\|_2\le
C_3\sum_{|j|>N}e^{-cj^{q}}\log j$, the same exponential up to a logarithm.
The section's gap persists, $g_N\ge g/2$ for $N$ large, by standard Galerkin
spectral approximation for operators with discrete
spectrum~\cite{BabuskaOsborn1991} together with the monotone convergence of
\cite[Prop.~3.4]{CCM2025}; the residual bound then yields
$\|\xi^{(N)}-P_N\xi^{(\infty)}\|_2\le 2\|P_NQW_\lambda
Q_N\xi^{(\infty)}\|_2/g_N$ up to a normalization correction of order
$\|Q_N\xi^{(\infty)}\|_2^2$. Finally
$|\widehat T[e]|\le C_4\sqrt{2N{+}1}\,\|e\|_2$ on the disc, using the lattice
distance $d(\lambda^2,r)$. All polynomial and logarithmic factors are absorbed
into the $(1+o(1))$ in the exponent.
\end{proof}

\begin{corollary}\label{cor:single}
Under the hypotheses of Lemma~\ref{lem:tail},
$|\nu_1(N,\lambda^2)-\nu_1(\infty,\lambda^2)|\le\exp(-c\,N^{q}(1+o(1)))$: the
coefficient tail alone forces the zero displacement at the single exponent,
by Lemma~\ref{lem:tail} followed by Lemma~\ref{lem:rouche}.
\end{corollary}

The open content of the archimedean rate is thereby confined to the following
statement, recorded in three clauses that separate the objects involved: the
eigenvector tail of the full form, the archimedean identification of its rate,
and the doubling beyond Corollary~\ref{cor:single}.

\begin{conjecture}[Coefficient decay and zero displacement]\label{conj:decay}
Fix $\lambda^2$, let $q=q(\lambda^2)$ be the exponent of
Section~\ref{ssec:prime}, and let $\xi^{(\infty)}$ denote the limit
eigenvector at cutoff $\lambda^2$.
\begin{enumerate}
\item[(i)] \emph{(Gevrey tail, full form.)} There exist $c(\lambda^2)>0$ and
  $C$ with
  \[
    |\xi^{(\infty)}_k|\;\le\;C\,\exp\!\bigl(-c(\lambda^2)\,k^{q}\bigr).
  \]
\item[(ii)] \emph{(Leading rate.)} $c(\lambda^2)\,L/(2\pi)\to\delta=\pi/4$
  as $\lambda^2\to\infty$, with $\delta$ the $\Gamma_\mathbb{R}$ envelope
  rate of Section~\ref{ssec:arch}. The approach is from below: at finite
  cutoff the small-$k$ rate sits under $\delta$ and the deficit shrinks as
  the cutoff grows.
\item[(iii)] \emph{(Doubling at the zero.)} The first-zero displacement obeys
  \[
    |\nu_1(N,\lambda^2)-\nu_1(\infty,\lambda^2)|
    \;\le\;C\,\exp\!\bigl(-2\,c(\lambda^2)\,N^{q}(1+o(1))\bigr),
  \]
  twice the tail exponent of (i).
\end{enumerate}
\end{conjecture}

Three remarks delimit the clauses. Clause (i) is the Gevrey membership of
Section~\ref{ssec:prime} read per coefficient. Clauses (ii) and (iii) together
recover the parameter-free slope~\eqref{eq:slope} at $q\approx1$, the
prediction validated to $3\%$ at $\lambda^2=7$; the finite-$\lambda^2$
coefficient rate ($\approx0.71$ at $\lambda^2=100$, against
$\delta\approx0.785$) is then a data point for the finite-cutoff deficit
that clause (ii) asserts, not a discrepancy. The author's coefficient-level
measurements across $\lambda^2=13$--$200$ show the small-$k$ rate climbing
toward $\pi/4$ as the cutoff grows, with the mid-band rate below it. What
controls the full profile is open and under study. Clause (iii) carries the doubling, and its
open content is now exactly the factor $2$: by Lemma~\ref{lem:tail} and
Corollary~\ref{cor:single}, clause (i) already implies the displacement bound
at the single exponent $c(\lambda^2)N^{q}$, rigorously, while the observed
displacement decays at twice it. The mechanism, whether first-order
smallness of the transform difference at the zero itself or the second-order
eigenvalue identity of Prop.~\ref{prop:stretched}(ii) transported through the
spectrum--zeros identification, is part of what must be proven. None of the
clauses is a citation target: no classical result supplies the
$\Gamma_\mathbb{R}$ envelope for the finite ground-state transform (the
realization step, in which the finite transform inherits the completed
factor), so this is new analysis, precisely posed. \emph{Status: T2 modulo
Conjecture~\ref{conj:decay}, the tier carried by the displacement transfer
Lemma~\ref{lem:rouche} (T1); the open item is typed new analysis in
Table~\ref{tab:tiers}, with clause (ii) the specific home of $\delta=\pi/4$
and clause (i) the Gevrey membership shared with the exponent row.}

\subsection{The prime correction: $q(\lambda^2)$}\label{ssec:prime}

The leading rate gives $q=1$. The finite Euler product
$\prod_{p\le\lambda^2}(1-p^{-s})^{-1}$ modulates the archimedean envelope by an
oscillatory prime term $S(z)=\sum_{p\le\lambda^2}p^{-1/2}\cos(z\ln p)+\cdots$
of mean-square strength
\[
  \langle S^2\rangle=\tfrac12\sum_{p\le\lambda^2}p^{-1}
  =\tfrac12(\ln\ln\lambda^2+M)+o(1)
\]
by Mertens' theorem ($M$ the Mertens constant); equivalently the full
log-product is $\sum_{p\le\lambda^2}-\ln(1-1/p)=\gamma+\ln\ln\lambda^2$. More
primes roughen the eigenvector and lower the analyticity exponent below $1$.
Modeling the reciprocal exponent as a baseline plus the prime fluctuation,
\begin{equation}\label{eq:q}
  \frac1q = c_0+c_1\ln\ln\lambda^2,\qquad
  c_0=\gamma\ \text{(archimedean baseline)},\quad
  c_1=\tfrac12\ \text{(variance prefactor)},
\end{equation}
which gives $q(\lambda^2)=2/(2\gamma+\ln\ln\lambda^2)$. The numerator $2$ is
$1/c_1$, the reciprocal of the $\tfrac12$ in the mean square; the $\ln\ln$ is
Mertens.

We test~\eqref{eq:q} against per-$\lambda^2$ fits of the full
form~\eqref{eq:nmodes} (Table~\ref{tab:q}). A linear regression
$1/q=c_0+c_1\ln\ln\lambda^2$ over $\lambda^2\in\{7,13,20,25,50,100,150\}$ yields
\[
  c_0=0.556\ (\gamma=0.577),\qquad c_1=0.478\text{--}0.508\ (\tfrac12=0.500),
  \qquad R^2=0.98 .
\]
Euler--Mascheroni $\gamma$ appears as the intercept of a regression that does
not assume it; the $c_1$ interval spans the regression variants used. The
offset of the parameter-free curve is systematic, not centered scatter.
Taking $c_0=\gamma$ exactly,
the idealized $q=2/(2\gamma+\ln\ln\lambda^2)$ runs $5$--$7\%$ \emph{below} the
per-$\lambda^2$ fits across the whole range (e.g.\ $1.10$ vs $1.185$ at
$\lambda^2=7$; $0.75$ vs $0.785$ at $\lambda^2=100$), and the regression
intercept $0.556$ sits $4\%$ under $\gamma=0.577$, always on the same side. We
therefore report both curves side by side --- the fitted
$(c_0,c_1)=(0.556,\,0.478\text{--}0.508)$ and the idealized
$(\gamma,\tfrac12)$ --- and the identification of the intercept with $\gamma$
is a structural hypothesis the data supports to that one-signed accuracy, not
a measurement centered on it. \emph{Status: T3; the remaining lemma is a rigorous
variance-to-exponent map, now precisely posed on the Gevrey scale
(Section~\ref{ssec:stretch}): prove that the almost-periodic prime multiplier
$S$, of mean square $\tfrac12\sum_{p\le\lambda^2}p^{-1}$, degrades the Gevrey
index of the strip-analytic ($G^1$) archimedean part linearly in that mean
square --- i.e.\ $\xi\in G^{\mu}$ sharply with
$\mu=\gamma+\tfrac12\ln\ln\lambda^2$. The characterization of $G^{\mu}$ by
coefficient decay $\exp(-c|k|^{1/\mu})$~\cite{Rodino1993} makes this statement
equivalent to the measured tail; what is missing is the subordination
mechanism, a piece of new analysis rather than a citation.}

\begin{table}[t]
\centering
\caption{Per-$\lambda^2$ fits of~\eqref{eq:nmodes} (full $\tanh$ form,
mode-limited points). The onset constant $c$ clusters near $4/\pi=1.273$; the
exponent $q$ decreases through $1$ near $\lambda^2\approx13$, matching the
archimedean crossover. $\lambda^2=200$ (sparse, $8$ points) is shown but
excluded from the regression.}
\label{tab:q}
\begin{tabular}{@{}rccccccc r@{}}
\toprule
$\lambda^2$ & 7 & 13 & 20 & 25 & 50 & 100 & 150 & 200 \\
\midrule
$q$ & 1.185 & 0.985 & 0.900 & 0.880 & 0.835 & 0.785 & 0.760 & 0.805 \\
$c$ & 1.20 & 1.25 & 1.28 & 1.29 & 1.31 & 1.28 & 1.26 & 1.31 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{The Shannon coordinate}\label{ssec:coord}

\begin{lemma}[Coordinate from ceiling over slope]\label{lem:coord}
The natural coordinate of the ramp is $N/(\lambda^2\ln\lambda^2)$. The scale at
which the leading ramp meets its ceiling is
\[
  N^\ast=\frac{D_{\mathrm{Primes}}}{\partial D/\partial N}
  =\frac{(4\pi/\ln10)\lambda^2}{\pi^2/(\ln10\,\ln\lambda^2)}
  =\frac{4}{\pi}\,\lambda^2\ln\lambda^2 .
\]
\end{lemma}

The $\lambda^2$ comes from the prime ceiling (Section~\ref{sec:primes}); the
$\ln\lambda^2$ from the inverse of the archimedean slope~\eqref{eq:slope}.
The combination $\lambda^2\ln\lambda^2$, the time--frequency Shannon number
of the box (Landau--Pollak~\cite{LandauPollak1962};
Slepian--Pollak~\cite{SlepianPollak1961}; Landau--Widom~\cite{LandauWidom1980}),
is therefore forced, not posited. (Measured saturation occurs at
$N/(\lambda^2\ln\lambda^2)\approx2.0$--$2.3$, larger than the linear value
$4/\pi$ because the $q<1$ ramp reaches its ceiling later than the initial-slope
line.) The form and the constant separate cleanly: writing
$N^\ast=\lambda^2\ln\lambda^2/\delta$, the grouping $\lambda^2\ln\lambda^2$ is
forced by the ceiling's $\lambda^2$ scaling and the basis interval
$L=\ln\lambda^2$ (Lemma~\ref{lem:sampling}) alone, independently of the archimedean constant, while
only the multiplier $4/\pi=1/\delta$ carries the rate $\delta$. That multiplier
is the data-pinned piece: the full-ramp onset constant clusters at $4/\pi$ across
the cutoffs of Table~\ref{tab:q} (mean $\approx1.27$). \emph{Status: the
coordinate form is T2 (ceiling-over-slope inputs); the multiplier $4/\pi$ is T3,
data-pinned.}

\section{The prime ceiling $D_{\mathrm{Primes}}$}\label{sec:primes}

The prime channel of the intrinsic law~\eqref{eq:main-intrinsic} is the accuracy
at infinite $N$ and $P$, set solely by which primes $p\le\lambda^2$ enter the
form. It is the relative error of the first zero at the cutoff,
\begin{equation}\label{eq:epsprimes}
  \varepsilon_{\mathrm{primes}}(\lambda^2)
  =\frac{|\nu_1(\infty,\lambda^2)-\gamma_1|}{|\gamma_1|},
\end{equation}
with $\nu_1(\infty,\lambda^2)$ the first zero at infinite basis size and fixed
$\lambda^2$: it is what remains once modes and precision are exhausted, and its
digit measurement $D_{\mathrm{Primes}}=-\log_{10}\varepsilon_{\mathrm{primes}}$ is
the ceiling the minimum~\eqref{eq:main} consumes. We fix its leading rate through
the plunge \emph{eigenvalue}, the tractable route, kept here in digit space to
reinforce the intrinsic statement with the measured quantity.

The plunge eigenvalue gives an \emph{eigenvalue floor}
$D_{\mathrm{eig}}=-\log_{10}\varepsilon^\infty(\lambda)$, the digit value of the
smallest eigenvalue. The CCM construction ties $\varepsilon^\infty$ to the
principal prolate spheroidal eigenvalue at bandwidth $\lambda$; the prolate
eigenvalue asymptotic (Connes~\cite{Connes2026}, \S6.4; Fuchs~\cite{Fuchs1964};
cf.\ Connes~\cite{ConnesProlate} and Hedenmalm~\cite{Hedenmalm2026}) gives, for
the relevant index,
\[
  1-\chi(\lambda)\sim \frac{2^{14}\sqrt{2}\,\pi^5}{3}\,
  \exp\!\bigl(-4\pi\lambda^2+\tfrac92\ln\lambda^2\bigr),
\]
so that
\begin{equation}\label{eq:eigfloor}
  D_{\mathrm{eig}}(\lambda^2)=K\lambda^2-\frac{9}{2\ln10}\ln\lambda^2-C_0,
  \qquad K=\frac{4\pi}{\ln10}=5.45750\ldots,
\end{equation}
with every constant closed-form: the leading $K=4\pi/\ln10$ is the base-$10$
rendering of the $e^{-4\pi\lambda^2}$ prolate decay rate that appears in Connes'
own framework, the $\tfrac92$ subleading-log coefficient is the Fuchs index, and
$C_0=\log_{10}\!\bigl(2^{14}\sqrt{2}\,\pi^5/3\bigr)=6.3736\ldots$ follows from the
prolate asymptotic (Groskin~\cite{Groskin2026} independently evaluates the same
\S6.4 prefactor and obtains $\log_{10}\approx6.37$, in agreement). This fixes the
leading rate of $\varepsilon_{\mathrm{primes}}$: the plunge sets $K\lambda^2$.

One direction of this rate can be anchored rigorously, and we record it as a
proposition. Recall that the bottom of the spectrum of the continuum form is a
variational infimum, $\varepsilon^\infty(\lambda)=\inf_\varphi
QW_\lambda(\varphi,\varphi)/\|\varphi\|^2$, so \emph{any} trial vector bounds it
from above, irrespective of the sign of $\varepsilon^\infty$.

\begin{proposition}[Variational one-sided floor bound]\label{prop:variational}
Let $k_\lambda$ be the prolate trial vector of the CCM prolate transfer
(\cite[\S\S7--8]{CCM2025}; \cite{ConnesConsaniZC}), built from the principal
prolate spheroidal function at bandwidth $\lambda$. Then, unconditionally,
\[
  \varepsilon^\infty(\lambda)\;\le\;
  R[k_\lambda]:=\frac{QW_\lambda(k_\lambda,k_\lambda)}{\|k_\lambda\|^2}.
\]
\end{proposition}
\begin{proof}
The bottom of the spectrum is the infimum of the Rayleigh quotient over the
form domain; any admissible trial vector bounds the infimum from above.
\end{proof}

\noindent The proposition is the variational principle verbatim, and is T1
with no caveat. The content is the size of the right-hand side, which we now
bound. Take $k_\lambda=\psi_0$, the principal prolate spheroidal mode at
time--bandwidth $c=2\pi\lambda^2$ transferred into the $V_n$ basis. On the
localized form's split $QW_\lambda=A_{\mathrm{arch}}-A_{\mathrm{prime}}$
(archimedean part $w_{02}-w_r$, prime part the prime-power sum $w_p$;
\cite[\S6]{CCM2025}) the quotient separates,
\begin{equation}\label{eq:Rsplit}
  R[k_\lambda]
   =\underbrace{\frac{\langle A_{\mathrm{arch}}\psi_0,\psi_0\rangle}{\|\psi_0\|^2}}
     _{=\,1-\chi_0(c)}
   -\underbrace{\frac{\langle A_{\mathrm{prime}}\psi_0,\psi_0\rangle}{\|\psi_0\|^2}}
     _{=\,\rho_p(\lambda)} ,
\end{equation}
where the archimedean term equals the prolate \emph{deficiency} $1-\chi_0(c)$
under the band-restricted identification of \cite[\S6.4]{CCM2025}, which we
take as a hypothesis alongside the sign condition below, and
$\rho_p(\lambda)$ is the prime Rayleigh quotient on $\psi_0$.

\begin{lemma}[Prolate floor bound]\label{lem:floorbound}
If the prime Rayleigh quotient on the principal prolate mode is non-negative,
$\rho_p(\lambda)\ge0$, then
\[
  \varepsilon^\infty(\lambda)\;\le\;1-\chi_0(2\pi\lambda^2).
\]
\end{lemma}
\begin{proof}
By Prop.~\ref{prop:variational} and~\eqref{eq:Rsplit},
$\varepsilon^\infty(\lambda)\le R[k_\lambda]=(1-\chi_0(c))-\rho_p(\lambda)
\le 1-\chi_0(c)$, the last step by $\rho_p\ge0$.
\end{proof}

\noindent The hypothesis $\rho_p\ge0$ is a single sign, far weaker than
evaluating $\rho_p$, and it is the natural one: the plunge decomposition
$\varepsilon^\infty=(1-\chi_0)-\rho_p$ with $\varepsilon^\infty>0$ already
forces $\rho_p\le1-\chi_0$, and the paper's measured prime contributions on the
ground state are positive and $O(1)$ (Section~\ref{sec:discussion}). With the
Fuchs asymptotic~\cite{Fuchs1964} for the principal deficiency,
\[
  1-\chi_0(2\pi\lambda^2)
   =\tfrac{2^{14}}{3}\sqrt2\,\pi^5\,
     \exp\!\bigl(-4\pi\lambda^2+\tfrac92\ln\lambda^2\bigr)(1+o(1)),
\]
Lemma~\ref{lem:floorbound} yields the one-sided floor bound
\[
  D_{\mathrm{eig}}(\lambda^2)\;\ge\;K\lambda^2-\frac{9}{2\ln10}\ln\lambda^2
  -C_0 ,\qquad K=\frac{4\pi}{\ln10},\quad
  C_0=\log_{10}\!\bigl(\tfrac{2^{14}}{3}\sqrt2\,\pi^5\bigr),
\]
whenever $\varepsilon^\infty(\lambda)>0$. This is now a proof, not a stated
expectation: the leading coefficient $K$ and the subleading Fuchs slope
$\tfrac92$ are \emph{rigorously bounded from below} on the floor, modulo only
the sign hypothesis $\rho_p\ge0$ (checkable directly through the archimedean/
prime Rayleigh split of \cite[\S6]{CCM2025}) and the classical Fuchs asymptotic.
The reverse inequality,
that the ceiling cannot be beaten, i.e.\ a \emph{lower} bound on
$\varepsilon^\infty$ and on $|\nu_1(\infty,\lambda^2)-\gamma_1|$ of the same
exponential order, remains open, and is the archimedean--prime cancellation
of Section~\ref{sec:discussion}.

Where finite-$\lambda^2$ control (rather than asymptotics) is needed, for
instance the curvature that produces the empirically steeper matching-floor
slope $P_m\approx5$ below, the modern non-asymptotic prolate literature
supplies two-sided eigenvalue bounds with explicit constants: the
three-region estimates of Bonami--Jaming--Karoui~\cite{BJK2021}, the improved
bounds of Karnik--Romberg--Davenport~\cite{KRD2021}, the upper bounds of
Osipov~\cite{Osipov2013}, the super-exponential decay estimates of
Bonami--Karoui~\cite{BonamiKaroui2017}, and Kulikov's exponential lower bound
before the plunge region~\cite{Kulikov2023}. These replace the
asymptotic-only inputs wherever a quantified inequality at a specific cutoff
is wanted.

The eigenvalue depth is a scaffold, not the object. What the law consumes is the
zero error $\varepsilon_{\mathrm{primes}}$, which sits a few digits below the
eigenvalue floor: converting the plunge depth $\varepsilon^\infty$ into the zero
displacement $|\nu_1-\gamma_1|$, and normalizing by $|\gamma_1|$ (the
$\log_{10}|\gamma_1|\approx1.15$ digits of the relative error), gives the matching
floor
\begin{equation}\label{eq:primes}
  D_{\mathrm{Primes}}(\lambda^2)=K\lambda^2-P_m\log_{10}\lambda^2-c_m,
  \qquad P_m\approx5,\quad c_m\approx9.5.
\end{equation}
The leading $K\lambda^2$ is shared with the eigenvalue floor, the same plunge
rate, so the intrinsic content is a \emph{single} channel, not two competing
ceilings; the subleading $P_m,c_m$ are the eigenvalue-to-zero conversion together
with the $|\gamma_1|$ normalization. Equivalently the \emph{conversion gap}
$\mathrm{conv}(\lambda^2):=D_{\mathrm{eig}}-D_{\mathrm{Primes}}
=(P_m-\tfrac92)\log_{10}\lambda^2+(c_m-C_0)\approx\tfrac12\log_{10}\lambda^2+3$
is that conversion, read in digit space. The leading rate $K$ is derived (prolate
transfer, one side now anchored by Prop.~\ref{prop:variational}); the
conversion, hence $P_m$ and $c_m$, is empirical; of it, the
$\approx1.15$-digit part of $c_m$ is the explicit $|\gamma_1|$ normalization and
the remainder is the depth-to-zero factor. A structural route to that factor exists, and we record it as a proposal
rather than claim it as a derivation. The zero-displacement identity of
Lemma~\ref{lem:rouche}, applied at the cutoff, gives
$|\nu_1(\infty,\lambda^2)-\gamma_1|\le
2\,\eta_\lambda/|\widehat\xi'(\gamma_1)|$, with $\eta_\lambda$ the
plunge-scale perturbation of the transform. The numerator carries the plunge
depth through the determinant identity
$\det{}_{\!\mathrm{reg}}(D-z)=-i\lambda^{-iz}\widehat\xi(z)$ of
\cite[Thm.~5.10]{CCM2025}; the denominator carries the $L$-dependent
normalization of the transform and the local zero density $\rho_Q(\gamma_1)$
of Section~\ref{sec:equilibrium}. Deriving the $\lambda$-scaling of that ratio
would promote $(P_m,c_m)$ from empirical to structural; until that is done the
row remains T4, with the route named.

The arithmetic does not move the eigenvalue-floor slope. At the scale the
operator resolves, the prime part of the Weil form has large-$|t|$ behavior
$S(t)=\tfrac12 t^2-\gamma|t|+c_0$ (Mertens' theorems with prime-number-theorem
partial summation). On the even, mean-zero ground eigenvector each piece is inert
on the prefactor power: $\tfrac12 t^2$ acts as a rank-one \emph{odd} operator that
annihilates the even ground state; $-\gamma|t|$ equals $2\gamma K_a$, a multiple of
the metric operator $K_a=(-\Delta)^{-1}$, so it only rescales the eigenvalue and
shifts the additive constant; and $c_0$ projects out of the mean-zero subspace
(the screw-function operator split of~\cite{Suzuki2026}). None of the leading
prime terms can therefore supply a $(\lambda^2)^\Delta$ prefactor correction:
given the rate, the prefactor power is the Fuchs index $\tfrac92$ of the prolate
model (Section~\ref{sec:primes}), and the entire arithmetic contribution is
confined to the additive constant. This inertness is a statement about the
prefactor only; the archimedean part of the form does not by itself carry the
plunge (it is $O(1)$-indefinite; see the rate-origin discussion in
Section~\ref{sec:discussion}). The empirically steeper
matching-floor slope $P_m\approx5$ is a finite-$\lambda^2$ curvature effect, not a
change in the asymptotic slope.

\emph{Status: the leading coefficient $K$ and the eigenvalue-floor slope $\tfrac92$
ride the prolate transfer (T2), with one side anchored by the variational bound
of Prop.~\ref{prop:variational} (the named lemma being the trace-formula
evaluation of $R[k_\lambda]$) and the reverse inequality open (new analysis:
the archimedean--prime cancellation); the arithmetic is provably inert on the
slope; only the matching-floor offset ($P_m$, $c_m$), the eigenvalue-to-zero
conversion, is empirical (T4), with a structural route named above.}

\paragraph{The floor is smooth, not a prime staircase.}
A natural hypothesis, that $D_{\mathrm{Primes}}$ steps at prime entries,
is falsified directly. With $N$ and $P$ held above floor ($N=300$, $P=200$),
sweeping $\lambda^2$ in unit steps from $4$ to $30$ gives a smooth rise of
$\approx5.3$ digits per unit $\lambda^2$; the per-unit rate at prime-power
entries is statistically indistinguishable from non-prime entries. The ceiling
is a smooth function of $\lambda^2$, consistent with the analytic
form~\eqref{eq:primes}.

\paragraph{Floor anchors.}
Floor-isolated measurements (both $N$ and $P$ above floor) give
$D_{\mathrm{Primes}}(100)=526.08$ at $(N,P)=(1000,700)$ and a saturated
$D_{\mathrm{Primes}}(200)\approx1070.4$ at $N\ge2400$. The latter sits at the
matching floor $K\cdot200-5\log_{10}200-9.5=1070.5$, confirming the two-floor
structure at the largest cutoff measured. Regarding provenance, the $\lambda^2=200$
anchor was measured against the $2500$-digit Arb reference of
Definition~\ref{def:D}, not a shorter tabulation. A matching-digit count cannot exceed
the length of its reference: against a $\sim1000$-digit reference the
measurement would \emph{saturate} near $999$ and could never read
$1070.4$, so the observed value is itself incompatible with a
reference-limited artifact. Such an artifact would present as a suspicious
plateau at the reference length, not a reading $70$ digits past it.

\section{The equilibrium-measure edge}\label{sec:equilibrium}

Sections~\ref{sec:nmodes} and~\ref{sec:primes} derived the mode ramp and the
prime ceiling by two separate classical routes: the Landau--Widom transition
width for the ramp, and the Fuchs prolate tail for the ceiling. This section
states the structural fact that ties them together --- both are features of a
single spectral edge, so the intrinsic law~\eqref{eq:main-intrinsic} is one
edge rather than two coincident mechanisms. It adds no new constant; it is the
reason the two budgets share a form.

\subsection{The archimedean equilibrium measure}\label{ssec:eqmeasure}

Treat the limiting ordinates $\{\nu_k\}$ as a unit charge in the external field
set by the archimedean weight of the Weil form. As $\lambda^2\to\infty$ the
macroscopic distribution is the equilibrium measure $\mu_Q$ minimizing the
weighted logarithmic energy
\[
  I[\mu]=\iint\log\frac{1}{|s-t|}\,d\mu(s)\,d\mu(t)+2\int Q(t)\,d\mu(t),
  \qquad \mu\ge0,\ \textstyle\int d\mu=1,
\]
with external field $Q$ the archimedean potential whose derivative is the phase
rate $Q'(t)=\tfrac12\log(t/2\pi)$ of Section~\ref{ssec:arch}. That a limiting
spectral or zero density is the equilibrium measure of such a weighted energy
problem is classical (Saff--Totik~\cite{SaffTotik1997}; in the
orthogonal-polynomial and random-matrix setting, Deift~\cite{Deift1999}); we
invoke it here as a known tool. The Euler--Lagrange condition
$2\int\log|s-t|\,d\mu(t)=2Q(s)+\mathrm{const}$ on $\operatorname{supp}\mu$ fixes
the density; with this field it is
\begin{equation}\label{eq:rvm}
  \rho_Q(t)=\frac{1}{2\pi}\log\frac{t}{2\pi},
\end{equation}
the Riemann--von Mangoldt zero-counting density
($\int_0^T\rho_Q=\tfrac{T}{2\pi}\log\tfrac{T}{2\pi e}$). The smooth backbone of
the spectrum is thus the equilibrium measure of the archimedean field, recovered
without reference to the primes and independent of RH.

The variational identification can be re-based on Szeg\H{o} theory, which turns
the open step into a \emph{named} lemma. The archimedean part of the truncated
matrix is a Toeplitz-type operator whose symbol is explicit, the phase rate
$\partial_t\theta(t)=\tfrac12\log(t/2\pi)+O(t^{-2})$ recorded in
\cite[eq.~(3.24)]{CCM2025}, and first-order Szeg\H{o} /
Kac--Murdock--Szeg\H{o} asymptotics for operators of this symbol
class~\cite{KMS1953,Widom1982} yield the bulk eigenvalue-counting law directly,
which is precisely the density~\eqref{eq:rvm}. The arithmetic part is a
\emph{bounded} perturbation uniformly in $N$ (the content of the discreteness
argument in \cite[Thm.~3.6]{CCM2025}: the contribution of the non-archimedean
places to the operator is bounded), so Weyl's inequality transfers the bulk law
to the full form unchanged up to an $O(1)$ shift. This two-step route ---
first-order Szeg\H{o} for the log symbol, plus bounded perturbation ---
replaces the variational analogy with checkable classical inputs; the
second-order term of the same theory
\cite{LandauWidom1980,Widom1982,Israel2015} is then the transition width of
Section~\ref{ssec:oneedge}. \emph{Status: T2 --- the density~\eqref{eq:rvm} is
the standard archimedean main term, the bounded-perturbation step is proven
in~\cite{CCM2025}, and the remaining open step is the named first-order
Szeg\H{o} lemma for this symbol class applied to the truncated operator;
the tier carried by the Szeg\H{o} route above.}

\subsection{One edge, two scales}\label{ssec:oneedge}

The equilibrium measure has a lower band edge: the eigenvalues descend from the
$O(1)$ bulk into the deep, positive plunge (positivity forced by the Weil form).
That edge is a single object, and the two budgets are its two scales:
\begin{itemize}
\item \emph{the transition is the mode ramp} --- the width of the crossover from
  bulk to plunge, as the resolved dimension crosses the Shannon number
  $\lambda^2\ln\lambda^2$, is the Landau--Widom log-width~\cite{LandauWidom1980},
  which is the $\tanh$ ramp of $D_{\mathrm{nModes}}$ (Section~\ref{sec:nmodes});
\item \emph{the extreme tail is the prime ceiling} --- the depth past the edge is
  the Fuchs extreme-value asymptotic~\cite{Fuchs1964}, which is
  $D_{\mathrm{Primes}}$ (Section~\ref{sec:primes}).
\end{itemize}
So the Landau--Widom width and the Fuchs tail, invoked separately above, are the
transition and the tail of the \emph{same} band edge. This is the mechanism
behind the algebra $D_{\mathrm{nModes}}=D_{\mathrm{Primes}}\cdot g$ of
Section~\ref{ssec:twoforms}: $D_{\mathrm{nModes}}$ and $D_{\mathrm{Primes}}$ are
one edge $D_{\mathrm{edge}}$ read at two resolutions, and the intrinsic
error~\eqref{eq:main-intrinsic} is correspondingly one edge: its
finite-$N$ transition $\varepsilon_{\mathrm{modes}}$ and its extreme tail
$\varepsilon_{\mathrm{primes}}$ are the ramp and the ceiling of $\mu_Q$'s lower
edge. The unification is conceptual, not a new estimate: the microscopic
universality class of the edge (Airy, Bessel, or a Landau--Widom band edge) and
the exact deep-edge constant remain open, the latter being the same
archimedean--prime cancellation left open for $D_{\mathrm{Primes}}$ in
Section~\ref{sec:primes}. The equilibrium-measure framework itself is classical
(above); to the best of the author's knowledge, its application to the
Connes--Consani--Moscovici / Connes--van Suijlekom Weil form (identifying the
zero-counting backbone as the archimedean equilibrium measure, and the mode ramp
and prime ceiling as the transition and extreme tail of that measure's lower
edge) is new here. \emph{Status: T3 --- structural unification riding the
prolate transfer already used for both budgets; the bulk beneath it is now T2
via the Szeg\H{o} route of Section~\ref{ssec:eqmeasure}, while the edge
universality class and the deep-edge constant remain open (new analysis).}

\section{The complete formula}\label{sec:complete}

\subsection{The assembled law}\label{ssec:assembled}

Collecting Sections~\ref{sec:wprec}--\ref{sec:equilibrium}, the law stands in the
two forms of the introduction. \emph{Intrinsically} it is the single edge of
Section~\ref{sec:equilibrium}: the relative error obeys the two-channel
decomposition
\begin{equation}\label{eq:complete-intrinsic}
  E(N,\lambda^2)\le\varepsilon_{\mathrm{modes}}(N,\lambda^2)
                +\varepsilon_{\mathrm{primes}}(\lambda^2)
\end{equation}
(an equality up to sign cancellation; Prop.~\ref{prop:min}); equivalently, in
digit units, the edge function
$D_{\mathrm{edge}}(N,\lambda^2)=D_{\mathrm{Primes}}(\lambda^2)\cdot
\tanh\!\bigl((c\,N/(\lambda^2\ln\lambda^2))^{q}\bigr)$, whose $N\to\infty$ ceiling
$D_{\mathrm{edge}}(\infty,\lambda^2)=D_{\mathrm{Primes}}$ is the prime floor and
whose finite-$N$ transition is the mode ramp $D_{\mathrm{nModes}}$.
\emph{Observed} at working precision $P$, it is that one edge capped by the
instrument resolution:
\begin{equation}\label{eq:complete}
  D(\lambda^2,N,P)=\min\!\Bigl(
  \underbrace{P+O(1)}_{D_{\mathrm{Wprec}}},\;
  \underbrace{D_{\mathrm{Primes}}\tanh\!\bigl((c\,N/(\lambda^2\ln\lambda^2))^{q}\bigr)}_{D_{\mathrm{nModes}}\;=\;D_{\mathrm{edge}}},\;
  \underbrace{K\lambda^2-P_m\log_{10}\lambda^2-c_m}_{D_{\mathrm{Primes}}\;=\;D_{\mathrm{edge}}(\infty)}
  \Bigr),
\end{equation}
with $K=4\pi/\ln10$ exact, $q=2/(2\gamma+\ln\ln\lambda^2)$, and $c\approx4/\pi$.
The last two entries are the one edge $D_{\mathrm{edge}}$ at finite $N$ and at
$N\to\infty$, so the operative law is $D=\min(D_{\mathrm{Wprec}},
D_{\mathrm{edge}})+O(1)$: a single spectral edge, capped by precision. This is
the Andrews CCM Convergence Law. Of its parts, the minimum bound, the
precision slope, the stretched-exponential rate unification
(Prop.~\ref{prop:stretched}), and the zero-displacement transfer
(Lemma~\ref{lem:rouche}) are rigorous; every remaining component carries the
explicit tier, open item, and open-item type recorded in
Table~\ref{tab:tiers}, from the T2 constants through the T3 exponent and onset
constant to the single empirical offset (T4).

To the best of the author's knowledge, the complete convergence model
of~\eqref{eq:complete} is new: the assembly of the three budgets into a single law, the closed-form
mode exponent $q(\lambda^2)=2/(2\gamma+\ln\ln\lambda^2)$, the Shannon coordinate
$\lambda^2\ln\lambda^2$, the stretched-exponential identification of the mode
budget, and the recognition of the mode ramp and prime ceiling as one
equilibrium-measure edge appear here for the first time. The rate it characterizes is precisely the
problem named open in~\cite{CCM2025}: Groskin~\cite{Groskin2026} measures a
Galerkin exponent but disclaims deriving it, and \'Sliwi\'nski~\cite{Sliwinski2026}
establishes only a lower bound, so no prior work supplies the functional form
given here. The novelty claimed is the law as a whole and the identification of
its mechanism; the underlying construction~\cite{CCM2025,CvS2025}, the
Paley--Wiener strip mechanism~\cite{Groskin2026}, and the classical prolate and
Mertens inputs are credited to their sources throughout.

\subsection{Convergence criterion}\label{ssec:criterion}

\begin{proposition}[Convergence criterion and the prime bottleneck]\label{prop:limit}
Let $\nu_1=\nu_1(\lambda^2,N,P)$ and $D=-\log_{10}(|\nu_1-\gamma_1|/|\gamma_1|)$.
Assume, at each fixed cutoff considered, the finite-cutoff hypothesis
\begin{equation}\label{eq:poshyp}
  (\mathrm{H}_\lambda):\qquad \varepsilon^\infty(\lambda)>0
  \quad\text{and}\quad \nu_1(\infty,\lambda^2)\neq\gamma_1 .
\end{equation}
\begin{enumerate}
\item[(i)] \emph{(Prime bottleneck.)} For every fixed $\lambda^2$ satisfying
  $(\mathrm{H}_\lambda)$,
  $\sup_{N,P}D(\lambda^2,N,P)=D_{\mathrm{Primes}}(\lambda^2)<\infty$. Hence no
  sequence with bounded $\lambda^2$ has $\nu_1\to\gamma_1$: convergence to the
  zero forces $\lambda^2\to\infty$.
\item[(ii)] \emph{(Criterion.)} Since $D_{\mathrm{nModes}}\le D_{\mathrm{Primes}}$
  (a mode-limited count cannot beat the ceiling) $D\to\infty$ if and only if
  $D_{\mathrm{Wprec}}\to\infty$ and $D_{\mathrm{nModes}}\to\infty$; the latter
  forces $\lambda^2\to\infty$ (part~(i)), with $N$ past the Shannon scale
  $\lambda^2\ln\lambda^2$ and $P\to\infty$. The mode budget rises monotonically
  toward its ceiling as $N\to\infty$ \cite[Prop.~3.4]{CCM2025}.
\end{enumerate}
Consequently, conditional on the continuum plunge
$D_{\mathrm{Primes}}(\lambda^2)\to\infty$ (i.e.\ $\varepsilon^\infty(\lambda)\to0$),
convergence holds along any path $\lambda^2\to\infty$ with
$N\gtrsim(4/\pi)\lambda^2\ln\lambda^2$ and $P\gtrsim K\lambda^2$, at rate
$D\sim K\lambda^2$.
\end{proposition}
\begin{proof}
(i) As $N\to\infty$ the truncation converges to the continuum operator at cutoff
$\lambda$, whose first zero lies at relative distance
$10^{-D_{\mathrm{Primes}}(\lambda^2)}$ from $\gamma_1$; here
$D_{\mathrm{Primes}}(\lambda^2)<\infty$ by hypothesis
$(\mathrm{H}_\lambda)$ and is $N$-independent
\cite[Thm.~1.3]{Suzuki2026}, and finite working precision only lowers $D$.
Hence $D\le D_{\mathrm{Primes}}(\lambda^2)$, finite, for all $N,P$.
(ii) By Prop.~\ref{prop:min}, $D=\min(D_{\mathrm{Wprec}},D_{\mathrm{nModes}},
D_{\mathrm{Primes}})+O(1)$; since $D_{\mathrm{nModes}}\le D_{\mathrm{Primes}}$,
this is $\min(D_{\mathrm{Wprec}},D_{\mathrm{nModes}})+O(1)$, so $D\to\infty$ if and
only if $D_{\mathrm{Wprec}}\to\infty$ and $D_{\mathrm{nModes}}\to\infty$; since
$D=-\log_{10}(|\nu_1-\gamma_1|/|\gamma_1|)$, this is exactly
$\nu_1\to\gamma_1$.
\end{proof}

\paragraph{Status of the hypothesis $(\mathrm{H}_\lambda)$.}
We state $(\mathrm{H}_\lambda)$ explicitly because finite-cutoff Weil
positivity is \emph{not} known unconditionally for all $\lambda$:
\cite{CCM2025} is explicit that positivity of the continuum measure cannot be
asserted in general, and positivity for all $\lambda$ is precisely the open
conjecture linking the construction to RH. What is unconditional is the
small-support range: Yoshida proved the Weil hermitian form positive definite
for test functions of sufficiently small support~\cite{Yoshida1992}, Bombieri
reproved and quantified this in his study of the truncated
functional~\cite{Bombieri2000}, Connes--Consani gave the operator-theoretic
strengthening at the archimedean place via the Sonin
space~\cite{ConnesConsaniSelecta}, and the $\zeta$-cycles
paper~\cite{ConnesConsaniZC} tracks finite-$\lambda$ positivity prime by prime
through the first cutoffs. Beyond that range $(\mathrm{H}_\lambda)$ is verified
numerically at every cutoff measured in this program ($\lambda^2\le200$, and
$\lambda^2\le1000$ in~\cite{AndrewsA}); the second clause
$\nu_1(\infty,\lambda^2)\neq\gamma_1$, that the truncated Euler product does
not accidentally reproduce the zero exactly, is likewise numerical, and is
what the finiteness of $D_{\mathrm{Primes}}$ actually uses.

\noindent This refines the qualitative monotone convergence of
\cite[Prop.~3.4]{CCM2025} into an explicit, conditional rate. The equivalence
$D\to\infty\Leftrightarrow\nu_1\to\gamma_1$ is definitional and rigorous (T1);
the prime bottleneck~(i) is rigorous \emph{conditional on}
$(\mathrm{H}_\lambda)$, which is unconditional in the Yoshida--Bombieri
small-support range and numerical beyond it; the conditional rate inherits the
rigor of $D_{\mathrm{Primes}}$: the prolate transfer (T2, one side anchored
by Prop.~\ref{prop:variational}) and the continuum plunge it assumes.

Conceptually, the criterion isolates the prime cutoff as the sole hard
bottleneck: precision and basis size are freely improvable and impose no ceiling,
so reaching the exact zero is, at root, a statement about $\lambda^2\to\infty$ and
the growth of the prime content. The rate $D\sim K\lambda^2$ quantifies the cost
--- each additional digit demands a fixed increment $\ln10/(4\pi)$ in $\lambda^2$
--- turning the construction's qualitative convergence into a concrete, if
conditional, budget.

\section{Discussion}\label{sec:discussion}

\paragraph{Relation to Groskin~\cite{Groskin2026}.}
Groskin independently implemented the Connes--van Suijlekom matrix and measured
a Galerkin convergence exponent $s(c)$ growing with the cutoff, proposing a
Paley--Wiener analyticity-strip mechanism. The present work builds on that
mechanism: Prop.~\ref{prop:stretched}(ii) shows his $s(c)$ is the local slope of our
stretched-exponential, and the strip width is identified with the archimedean
$\Gamma_\mathbb{R}$ rate $\delta=\pi/4$. What is new here is the digit-space
formulation, the Shannon coordinate, the closed form for $q(\lambda^2)$, and the
embedding of the rate in the full three-budget minimum.

Groskin's revised analysis, extended to cutoff $\lambda^2=100$, is independent
corroboration of this two-regime structure. He finds that no single
smooth-convergence law fits his eigenvalue-space data and explicitly declines to
assert a convergence rate, reporting instead an unresolved tension between a
\emph{drifting} effective exponent and an apparent \emph{finite} convergence
limit. Both halves are accounted for here: Prop.~\ref{prop:stretched}(ii) explains
the drift (a fixed Sobolev index cannot persist, since the local exponent
$2s_{\mathrm{loc}}(N)=aqN^{q}$ grows with $N$), while the prime ceiling of
Section~\ref{sec:primes} supplies the finite limit $D_{\mathrm{Primes}}(\lambda^2)$.
What looks like an irreconcilable choice within a single-rate framework is simply
the pre-saturation ramp and the ceiling of the minimum law~\eqref{eq:main}, seen
in one dataset.

\paragraph{Relation to \'Sliwi\'nski~\cite{Sliwinski2026} and Suzuki~\cite{Suzuki2026}.}
\'Sliwi\'nski proves an inverse-logarithmic lower bound on the \emph{mean}
spectral error over the whole spectrum (Thm.~3.1), a different regime from the
first-zero convergence here, and consistent with it, since the mean is dominated
by the unconverged high-index eigenvalues while $\nu_1\to\gamma_1$ converges
super-exponentially. Suzuki's screw-function analysis supplies the archimedean
$\ln|z|$ multiplier and the prime-frequency support that underlie
Sections~\ref{ssec:arch}--\ref{ssec:prime}.

\paragraph{Origin of the leading rate: arithmetic, not archimedean.}
The prime ceiling is \emph{modeled} by an archimedean object, the prolate
eigenvalue of Section~\ref{sec:primes}, yet the exponential rate is not carried
by the archimedean part of the Weil form itself. Diagonalizing the
screw-function split~\cite{Suzuki2026} at high precision, the archimedean-only
localized form is $O(1)$-indefinite, with no exponentially small eigenvalue.
Restricting it to the Sonin subspace, the orthogonal complement of the
band-concentrated prolate modes, where Connes--Consani establish archimedean Weil
positivity~\cite{ConnesConsaniSelecta} (surveyed in~\cite{Connes2026}),
restores positivity, but only as an $O(1)$ gap:
the smallest Sonin eigenvalue is $2.26$, $2.28$, $2.57$ at $\lambda^2=5,8,13$,
essentially flat in $\lambda^2$, against a floor $K\lambda^2=27,44,71$. A
direct diagonalization of the archimedean-only form at $\lambda^2=13$--$50$
finds the same structure from the eigenvector side: the lowest even mode is a
delocalized Sonin-type state with an $O(1)$ eigenvalue, so the
band-concentrated plunge eigenvector exists only for the full form.
Archimedean positivity and the plunge are therefore \emph{decoupled}: the place
at infinity supplies an $O(1)$ spectral gap, while the exponentially small
$\varepsilon^\infty\sim e^{-4\pi\lambda^2}$ is produced by the cancellation
between the archimedean and prime parts of the full form. This locates the single
open analytic step behind the leading coefficient: a quantitative
archimedean--prime cancellation estimate on the even ground eigenvector, the
precise sense in which $K$ is~T2.

\paragraph{Practical use.}
The law is a parameter-selection guide: to target $D^\ast$ digits, pick
$\lambda^2\gtrsim D^\ast\ln10/(4\pi)=D^\ast/5.46$ so the ceiling clears
$D^\ast$; pick $N$ so the $\tanh$ ramp clears $D^\ast$ (a few times the Shannon
number $\lambda^2\ln\lambda^2$); and pick $P\gtrsim D^\ast$. Increasing any one
parameter past the binding minimum yields nothing.

\section{Rigor ledger}\label{sec:rigor}

Table~\ref{tab:tiers} collects the rigor status of every component of the law,
together with the open item that would promote each to the next level. A
fourth column types each open item: \emph{named lemma} means
the missing step is a precisely posed statement inside an existing classical
theory (variational principle, Rouch\'e, Szeg\H{o}, Gevrey characterization),
so that closing it is a matter of executing known machinery; \emph{new
analysis} means genuinely new estimates are required. Four rows sit in the
second category.

\begin{table}[H]
\centering
\caption{Rigor status of each component. T1: rigorous. T2: derived modulo one
named lemma. T3: derived in structure and validated against data, with a map not
yet rigorous. T4: empirical. Open-item type: \emph{named lemma} = precisely
posed inside existing theory; \emph{new analysis} = new estimates required.}
\label{tab:tiers}
{\footnotesize\setlength{\tabcolsep}{5pt}%
\begin{tabular}{@{}l l p{4.9cm} l@{}}
\toprule
Component & Status & Open item & Type \\
\midrule
Minimum structure (Prop.~\ref{prop:min}), lower bound & \textbf{T1} & --- & --- \\
Minimum structure, matching upper bound & \textbf{T1}$^\ast$ & cancellation excluded at seams & empirical check \\
$D_{\mathrm{Wprec}}=P+O(1)$, slope $1$ & \textbf{T1} & implementation offset & --- \\
Stretched-exp rate unification (Prop.~\ref{prop:stretched}) & \textbf{T1} & --- & --- \\
Zero-displacement transfer (Lem.~\ref{lem:rouche}) & \textbf{T1} & --- & --- \\
Tail-to-transform transfer (Lem.~\ref{lem:tail}, Cor.~\ref{cor:single}) & \textbf{T1} & --- & --- \\
Convergence criterion (Prop.~\ref{prop:limit}) & \textbf{T1}$\,|\,(\mathrm{H}_\lambda)$ & positivity beyond Yoshida range & new analysis \\
$D_{\mathrm{Primes}}$ leading $K$: floor lower bound & \textbf{T2} & prime sign $\rho_p\ge0$ (Lem.~\ref{lem:floorbound}) & named lemma \\
$D_{\mathrm{Primes}}$ leading $K$: ceiling unbeatable & open & archimedean--prime cancellation & new analysis \\
$D_{\mathrm{Primes}}$ floor slope $9/2$ & \textbf{T2} & Fuchs index (Lem.~\ref{lem:floorbound}); primes provably inert & named lemma \\
$D_{\mathrm{Primes}}$ matching offset ($P_m,c_m$) & \textbf{T4} & $\lambda$-scaling of $\eta_\lambda/|\widehat\xi'(\gamma_1)|$, route named & named lemma \\
Shannon coordinate $\lambda^2\ln\lambda^2$ & \textbf{T2} & ceiling/slope (Lem.~\ref{lem:coord}) & named lemma \\
Equilibrium-measure bulk (RvM density) & \textbf{T2} & first-order Szeg\H{o} for the log symbol & named lemma \\
Edge unification (one edge, two scales) & \textbf{T3} & edge universality class; deep-edge constant & new analysis \\
Archimedean rate $\delta=\pi/4$ & \textbf{T2} & Conj.~\ref{conj:decay}(ii) (Lem.~\ref{lem:rouche} closes the transfer) & new analysis \\
Exponent $q=2/(2\gamma+\ln\ln\lambda^2)$ & \textbf{T3} & Gevrey subordination (Conj.~\ref{conj:decay}(i)) & new analysis \\
Onset constant $c$ ($\sim 4/\pi$) & \textbf{T3} & exact rational & --- \\
\bottomrule
\end{tabular}}
\medskip

{\footnotesize $^\ast$Unconditional as a lower bound; the equality direction
holds absent cancellation between the signed channels
(Prop.~\ref{prop:min}), verified across the experimental program.
$\mathrm{T1}\,|\,(\mathrm{H}_\lambda)$: rigorous conditional on the
finite-cutoff hypothesis~\eqref{eq:poshyp}, unconditional in the
Yoshida--Bombieri small-support range~\cite{Yoshida1992,Bombieri2000}.}
\end{table}

\section{Conclusion}\label{sec:conclusion}

We have given a single law~\eqref{eq:main} for the convergence of the CCM
zeta spectral triple. In it the minimum structure is forced; the prime ceiling
has the parameter-free leading coefficient $4\pi/\ln10$, rigorously anchored
from one side by the variational bound of Prop.~\ref{prop:variational} and
open in the reverse inequality; and the mode budget, the convergence rate
named open in~\cite{CCM2025}, is the logarithm of a stretched-exponential
whose exponent is the eigenvector's analyticity class, derived end to end as
$q=2/(2\gamma+\ln\ln\lambda^2)$. The mode ramp and the prime ceiling are not two
mechanisms but one object, the lower edge of the archimedean equilibrium
measure whose bulk density is the Riemann--von Mangoldt law, seen at two
scales (Section~\ref{sec:equilibrium}). Read as a convergence statement, the law says
$\nu_1\to\gamma_1$ exactly when all three budgets diverge, giving an explicit
rate $D\sim(4\pi/\ln10)\lambda^2$ along $\lambda^2\to\infty$
(Prop.~\ref{prop:limit}). Each claim is stated with an explicit rigor
level. The rigorous core (the minimum bound, the precision slope, the
stretched-exponential convergence proposition~\ref{prop:stretched} (which unifies the
digit- and eigenvalue-space rates), and the zero-displacement and
tail-to-transform lemmas~\ref{lem:rouche} and~\ref{lem:tail}) stands on its
own; the variational bound of
Prop.~\ref{prop:variational} anchors one side of the prime floor; and the
remaining open items are typed in Table~\ref{tab:tiers}. Of these, the
\emph{named lemmas} (the prime-sign hypothesis $\rho_p\ge0$ behind the
floor bound of Lem.~\ref{lem:floorbound} and the first-order Szeg\H{o}
asymptotics for the log symbol)
are precisely posed inside classical theories and are the natural next
steps; the \emph{new-analysis} items (the coefficient-decay estimate for
the $\Gamma_\mathbb{R}$-weighted form (Conjecture~\ref{conj:decay}), the
Gevrey subordination behind $q$,
the archimedean--prime cancellation behind the unbeatable ceiling, and
positivity beyond the Yoshida range) are where the construction's genuinely
open mathematics lives.

\section{Explicit non-claims}\label{sec:nonclaims}

The per-component rigor of the law is the subject of Table~\ref{tab:tiers} and
is not repeated here. Beyond it, we are explicit about three boundaries.

\begin{itemize}
\item \textbf{No proof of the Riemann Hypothesis.} We characterize how the
  CCM/CvS approximation converges; the link from the construction to RH,
  continuum positivity of $QW_\lambda$ for all $\lambda$, is the open
  conjecture of~\cite{CCM2025,CvS2025,Connes2026}, and is neither assumed nor
  advanced here.
\item \textbf{We do not establish that the construction reproduces the zeta
  zeros.} That the finite operator's zeros converge to the true ordinates as
  $\lambda^2\to\infty$ is the construction's own open premise; we quantify the
  approach to $\gamma_1$ \emph{under} that premise (Prop.~\ref{prop:limit}) and
  do not prove the premise itself.
\item \textbf{Scope.} The law is stated and tested for the first zero $\gamma_1$
  of the Riemann zeta function. We make no claim about its constants for higher
  zeros, nor about extension to Dirichlet $L$-functions or other settings.
  The rate-law constants $(q,c)$ are pinned on $\lambda^2\le150$
  ($\lambda^2=200$, sparse at $8$ points, is shown but excluded from the $q$
  regression); their constancy at large $\lambda^2$, the regime in which the
  law becomes a convergence statement, is untested, and we make no claim
  that the fitted forms persist beyond the measured range.
\item \textbf{No unconditional finite-cutoff positivity.} Statements requiring
  $\varepsilon^\infty(\lambda)>0$ or $\nu_1(\infty,\lambda^2)\neq\gamma_1$ are
  conditioned on the hypothesis~$(\mathrm{H}_\lambda)$ of
  eq.~\eqref{eq:poshyp}, which is unconditional only in the small-support range
  of Yoshida~\cite{Yoshida1992} and Bombieri~\cite{Bombieri2000} and is
  verified numerically beyond it. We do not claim positivity of $QW_\lambda$
  for general $\lambda$.
\end{itemize}

\section*{Acknowledgments}
The author is deeply grateful to Professor Alain Connes (Coll\`ege de France
and IH\'ES) for his warm correspondence and support --- in particular for
reviewing the author's earlier reproduction~\cite{AndrewsA} and sharing it with
fellow researcher Akiva Groskin --- and for his kind reception of an early
draft of this work. The author also thanks Professor Henri Moscovici (The Ohio
State University) and Professor Nigel Higson (The Pennsylvania State
University) for correspondence and encouragement.

\section*{Statement on the use of AI tools}
In keeping with the arXiv policy on authors' use of generative AI language
tools, the author discloses that Anthropic's Claude AI model served as an aid
in preparing the exposition and the Rust implementation code, to specifications
set by the author. AI assistance was used to improve prose formatting and
readability. It was also used to gather research material, propose candidate
arguments, and validate calculations; all such output was reviewed by the
author. The author originated, directed, and conducted every computational
experiment presented here. The mathematical development was led by the author
with assistance from AI. The author bears full responsibility for the contents
of this manuscript.

\section*{Empirical testing and code availability}
The results above rest on an extensive empirical program: over $5{,}000$
high-precision CCM experiments across the cutoff, basis-size, and
working-precision axes, very likely the largest set of CCM calculations
assembled to date, especially at large configurations. The testing came first
and the analysis followed. Candidate rate laws were tested against the data and
discarded when they failed~\cite{AndrewsB}, and each failure narrowed the space
of forms the operator would tolerate. Isolating one parameter at a time for
study, with the other two driven out of the way, is how each budget's form was
read, and the breadth of the dataset ($\lambda^2=200$ at $N\ge2400$ here,
$\lambda^2=1000$ in the reproduction~\cite{AndrewsA}) is what made that
isolation possible. The extensive empirical testing was not decoration but the
heart of how the author mapped the operator's behavior.

The paper source and reproduction scripts are available at
\url{https://github.com/TeamXcelerator/ccm-convergence-rate}; the shared
high-precision library (Xcelerator Toolkit) is at
\url{https://github.com/TeamXcelerator/xcelerator-toolkit}. Both are
source-available in accordance with the licensing provided with each repository.

\appendix
\section{Self-contained structural facts}\label{app:selfcontained}

This appendix records the two structural facts about $\widehat\xi$ used in the
body, the closed form of the transform (Lemma~\ref{lem:sampling}) and
entirety/reality of its zeros, in a form that depends only on elementary
computation and on the self-contained proofs of~\cite{CCM2025}. Its purpose is
to make explicit that no statement in this paper routes through the
Carath\'eodory--Fej\'er machinery of~\cite{CvS2025}, which we cite as the
historical origin of the reality mechanism only.

\subsection{The closed form of the transform, from scratch}\label{app:closedform}

Write $t=\ln u\in[-L/2,L/2]$, so that
$V_n(u)=L^{-1/2}e^{2\pi i n(t+L/2)/L}=L^{-1/2}(-1)^n e^{2\pi i nt/L}$ and the
Fourier--Mellin transform is the Fourier integral over the interval. For a
single mode,
\begin{align*}
  \widehat V_n(z)&=\int_{-L/2}^{L/2}e^{izt}\,V_n(t)\,dt
  =L^{-1/2}(-1)^n\int_{-L/2}^{L/2}e^{i(z+2\pi n/L)t}\,dt\\
  &=L^{-1/2}(-1)^n\,\frac{2\sin\!\bigl((z+2\pi n/L)\,L/2\bigr)}{z+2\pi n/L}.
\end{align*}
Since $\sin(zL/2+\pi n)=(-1)^n\sin(zL/2)$, the sign cancels:
\[
  \widehat V_n(z)=\frac{2L^{-1/2}\sin(zL/2)}{z+2\pi n/L}.
\]
Summing over the eigenvector $\xi=\sum_{|n|\le N}\xi_n V_n$ and reindexing
$j=-n$ (using $\xi_{-j}=\xi_j$ for the even eigenvector, or simply relabeling)
gives
\[
  \widehat\xi_N(z)=2L^{-1/2}\sin(zL/2)\sum_{|j|\le N}\frac{\xi_j}{z-2\pi j/L},
\]
the closed form of Lemma~\ref{lem:sampling}, in agreement with
\cite[Prop.~5.9]{CCM2025}. The derivation is a one-line Fourier integral; no
external input is used. Entirety is immediate: each apparent pole
$z=2\pi j/L$ is cancelled by the zero of $\sin(zL/2)$ there (the removable-pole
computation in the proof of Lemma~\ref{lem:sampling}), so $\widehat\xi_N$ is
entire of exponential type $L/2$.

\subsection{Reality of the zeros: provenance}\label{app:reality}

The statement that, in the critical case, all zeros of $\widehat\xi_N$ are real
and coincide with the spectrum of the perturbed operator is
\cite[Thm.~5.10]{CCM2025}. We record why its proof is self-contained there.
The proof chain consists of: (a) \cite[Lem.~5.2]{CCM2025}, the observation that
the relevant matrices have the structure $\tau_{ij}=(b_i-b_j)/(i-j)$; (b)
\cite[Lem.~5.4]{CCM2025}, the elementary commutator identity
$DT-TD=|\beta\rangle\langle\eta|-|\eta\rangle\langle\beta|$ for such matrices,
with $D$ the diagonal mode operator; and (c) the conclusion of
\cite[Thm.~5.10]{CCM2025}: when $\varepsilon_N$ is the smallest eigenvalue with
even simple eigenvector, $T=QW^N_\lambda-\varepsilon_N\ge0$, the perturbed
operator is self-adjoint in the $T$-inner product, and the transform is
proportional to a regularized characteristic determinant,
$\det_{\mathrm{reg}}(D''-z)=-i\lambda^{-iz}\widehat\xi(z)$, whose zeros are
therefore real, being the characteristic polynomial of a self-adjoint matrix. Each
of (a)--(c) is proven within~\cite{CCM2025} by direct computation; the general
Carath\'eodory--Fej\'er extension of~\cite{CvS2025}, of which this is the
specialization, is not invoked at any step used here.

\subsection{The independent classical foundation}\label{app:foundation}

For completeness we record the pre-CvS lineage on which the truncated Weil
form itself rests: Weil's positivity criterion; Yoshida's unconditional
positivity for test functions of sufficiently small
support~\cite{Yoshida1992}; Bombieri's study of the quadratic functional
restricted to a finite interval, its minimizer in the $L^2$ unit ball, and the
eigenvalues of its finite truncations~\cite{Bombieri2000}; the
operator-theoretic strengthening of Yoshida's theorem at the archimedean place
by Connes--Consani~\cite{ConnesConsaniSelecta}; the prolate identification of
the small eigenvectors in the $\zeta$-cycles paper~\cite{ConnesConsaniZC}; and
Suzuki's screw-function framework~\cite{SuzukiJLMS,SuzukiHilbert,Suzuki2026},
which unifies Yoshida, Bombieri, and the Connes--Consani(--Moscovici) results
by means of continuous functions and de~Branges-space methods. The
construction studied in this paper can be founded entirely on this lineage
together with~\cite{CCM2025}.

\begin{thebibliography}{99}

\bibitem{CCM2025}
A.~Connes, C.~Consani, H.~Moscovici,
\emph{Zeta Spectral Triples}, arXiv:2511.22755 (2025).

\bibitem{CvS2025}
A.~Connes, W.~D.~van Suijlekom,
\emph{Quadratic Forms, Real Zeros and Echoes of the Spectral Action},
arXiv:2511.23257 (2025).

\bibitem{ConnesProlate}
A.~Connes,
\emph{Zeta zeros and prolate wave operators}, arXiv:2310.18423.

\bibitem{Connes2026}
A.~Connes,
\emph{The Riemann Hypothesis: Past, Present and a Letter Through Time},
arXiv:2602.04022 (2026).

\bibitem{Suzuki2026}
M.~Suzuki,
\emph{Weil's quadratic form via the screw function}, arXiv:2606.09096 (2026).

\bibitem{Hedenmalm2026}
H.~Hedenmalm,
\emph{Spectral Interpretation of Riemann Zeta Zeros}, arXiv:2606.17494 (2026).

\bibitem{Groskin2026}
A.~Groskin,
\emph{High-Precision Approximation of Riemann Zeros via the Truncated Weil
Form}, arXiv:2605.20224 (2026).

\bibitem{Sliwinski2026}
D.~\'Sliwi\'nski,
\emph{Spectral Analysis of the $D_{\log}^{(\lambda,N)}$ Operators},
arXiv:2601.12133 (2026).

\bibitem{Yoshida1992}
H.~Yoshida,
\emph{On Hermitian forms attached to zeta functions}, in \emph{Zeta Functions
in Geometry} (Tokyo, 1990), Adv. Stud. Pure Math. \textbf{21}, Kinokuniya,
Tokyo, 1992, pp.~281--325.

\bibitem{Bombieri2000}
E.~Bombieri,
\emph{Remarks on Weil's quadratic functional in the theory of prime numbers, I},
Atti Accad. Naz. Lincei Cl. Sci. Fis. Mat. Natur. Rend. Lincei (9) Mat. Appl.
\textbf{11} (2000), no.~3, 183--233.

\bibitem{ConnesConsaniSelecta}
A.~Connes, C.~Consani,
\emph{Weil positivity and trace formula: the archimedean place},
Selecta Math. (N.S.) \textbf{27} (2021), no.~4, Paper No.~77.

\bibitem{ConnesConsaniZC}
A.~Connes, C.~Consani,
\emph{Spectral triples and $\zeta$-cycles},
Enseign. Math. \textbf{69} (2023), no.~1--2, 93--148.

\bibitem{SuzukiJLMS}
M.~Suzuki,
\emph{Aspects of the screw function corresponding to the Riemann zeta function},
J. Lond. Math. Soc. (2) \textbf{108} (2023), no.~2.

\bibitem{SuzukiHilbert}
M.~Suzuki,
\emph{On the Hilbert space derived from the Weil distribution},
Canad. J. Math. (to appear); arXiv:2301.00421.

\bibitem{BJK2021}
A.~Bonami, P.~Jaming, A.~Karoui,
\emph{Non-asymptotic behavior of the spectrum of the sinc-kernel operator and
related applications}, J. Math. Phys. \textbf{62} (2021), 033511.

\bibitem{KRD2021}
S.~Karnik, J.~Romberg, M.~A.~Davenport,
\emph{Improved bounds for the eigenvalues of prolate spheroidal wave functions
and discrete prolate spheroidal sequences},
Appl. Comput. Harmon. Anal. \textbf{55} (2021), 97--128.

\bibitem{Osipov2013}
A.~Osipov,
\emph{Certain upper bounds on the eigenvalues associated with prolate
spheroidal wave functions},
Appl. Comput. Harmon. Anal. \textbf{35} (2013), 309--340.

\bibitem{BonamiKaroui2017}
A.~Bonami, A.~Karoui,
\emph{Spectral decay of time and frequency limiting operator},
Appl. Comput. Harmon. Anal. \textbf{42} (2017), 1--20.

\bibitem{Kulikov2023}
A.~Kulikov,
\emph{Exponential lower bound for the eigenvalues of the time-frequency
localization operator before the plunge region}, arXiv:2306.12430.

\bibitem{Israel2015}
A.~Israel,
\emph{The eigenvalue distribution of time-frequency localization operators},
arXiv:1502.04404.

\bibitem{KMS1953}
M.~Kac, W.~L.~Murdock, G.~Szeg\H{o},
\emph{On the eigenvalues of certain Hermitian forms},
J. Rational Mech. Anal. \textbf{2} (1953), 767--800.

\bibitem{Widom1982}
H.~Widom,
\emph{On a class of integral operators with discontinuous symbol},
in \emph{Toeplitz Centennial}, Operator Theory: Adv. Appl. \textbf{4},
Birkh\"auser, Basel, 1982, pp.~477--500.

\bibitem{Rodino1993}
L.~Rodino,
\emph{Linear Partial Differential Operators in Gevrey Spaces},
World Scientific, Singapore, 1993.

\bibitem{Fuchs1964}
W.~H.~J.~Fuchs,
\emph{On the eigenvalues of an integral equation arising in the theory of
band-limited signals}, J. Math. Anal. Appl. \textbf{9} (1964), 317--330.

\bibitem{LandauWidom1980}
H.~J.~Landau, H.~Widom,
\emph{Eigenvalue distribution of time and frequency limiting},
J. Math. Anal. Appl. \textbf{77} (1980), 469--481.

\bibitem{SlepianPollak1961}
D.~Slepian, H.~O.~Pollak,
\emph{Prolate spheroidal wave functions, Fourier analysis and uncertainty --- I},
Bell Syst. Tech. J. \textbf{40} (1961), 43--63.

\bibitem{LandauPollak1962}
H.~J.~Landau, H.~O.~Pollak,
\emph{Prolate spheroidal wave functions, Fourier analysis and uncertainty ---
III: the dimension of the space of essentially time- and band-limited signals},
Bell Syst. Tech. J. \textbf{41} (1962), 1295--1336.

\bibitem{SaffTotik1997}
E.~B.~Saff, V.~Totik,
\emph{Logarithmic Potentials with External Fields},
Grundlehren der mathematischen Wissenschaften \textbf{316}, Springer, 1997.

\bibitem{Deift1999}
P.~Deift,
\emph{Orthogonal Polynomials and Random Matrices: A Riemann--Hilbert Approach},
Courant Lecture Notes in Mathematics \textbf{3}, American Mathematical Society,
1999.

\bibitem{BabuskaOsborn1991}
I.~Babu\v{s}ka, J.~Osborn,
\emph{Eigenvalue problems}, in \emph{Handbook of Numerical Analysis}, Vol.~II
(Finite Element Methods, Part~1), P.~G.~Ciarlet and J.~L.~Lions, eds.,
North-Holland, 1991, pp.~641--787.

\bibitem{AndrewsA}
R.~Andrews, Jr.,
\emph{Independent Reproduction and Convergence Analysis of the
Connes--Consani--Moscovici Zeta Spectral Triple}, Zenodo (2026),
doi:10.5281/zenodo.20427499.

\bibitem{AndrewsB}
R.~Andrews, Jr.,
\emph{Empirical Falsification of Convergence Rate Predictions for the
Connes--Consani--Moscovici Zeta Spectral Triple}, Zenodo (2026),
doi:10.5281/zenodo.20427673.

\bibitem{Arb2017}
F.~Johansson,
\emph{Arb: efficient arbitrary-precision midpoint--radius interval arithmetic},
IEEE Trans. Comput. \textbf{66} (2017), no.~8, 1281--1292.

\bibitem{Odlyzko}
A.~M.~Odlyzko,
\emph{Tables of zeros of the Riemann zeta function},
\url{http://www.dtc.umn.edu/~odlyzko/zeta_tables/}.

\end{thebibliography}

\end{document}
